---
title: "Campaign Data Summary"
author: "Magali Blanco"
date: ' `r Sys.time()` '
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
    number_sections: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
editor_options: 
  chunk_output_type: console
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      cache=F, cache.comments = F, 
                      message = F, warning = F, 
                      tidy.opts=list(width.cutoff=60), tidy=TRUE,
                      #fig.height = 5, fig.width = 8
                      fig.height = 6, fig.width = 10
                      )  

# # Clear workspace of all objects and unload all extra (non-base) packages
# rm(list = ls(all = TRUE))
# if (!is.null(sessionInfo()$otherPkgs)) {
#   res <- suppressWarnings(
#     lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
#       detach, character.only=TRUE, unload=TRUE, force=TRUE))
# }

pacman::p_load(knitr, kableExtra, 
               ggpubr, tidyverse,
               ggrepel, #geom_label_repel
               ggmap, sf, ggspatial, #mapping...adding scales, N arrows
               ggrepel, #avoid overlapping labels
               units, #convert between e.g., m to km
               #time series data 
               lubridate
               )    
 

set.seed(1)

```


# Upload data

```{r}
#functions
source("0_Functions.R")

#only need location table
load(file.path("Data", "Original", "imports_fieldnotes_location_tables.rda"))
rm(imports, fieldnotes)

#definitions: ap 
load(file.path("Data", "Output", "common_vars.rda"))

#stop data
sec0 <- readRDS(file.path("Data", "Output", "second_data.rda"))
stops0 <- readRDS(file.path("Data", "Output", "stop_data.rda"))

```

```{r}
#mapping
crs_m <- 32148
crs_deg <- 4326 #WGS84. in decimal degrees

#shapefiles
monitoring_area_shp <- read_sf(file.path("..", "GIS", "Shapefiles", "Monitoring", "monitoring_area_filled_in.shp")) %>%
  #convert from 4269 
  st_transform(crs_deg)

```

### --> calcualted PM2.5 from neph. do this for sec data as well?

```{r}
sec <- sec0 %>%
  mutate(
    
    #use doe neph (bscat x10^4 values) and pm2.5 calibration curve to calc pm2.5 from our neph readings
    value = ifelse(grepl("neph", variable), value*10^4*doe_slope + doe_int, value),
    variable = ifelse(grepl("neph", variable), "pm2.5_ug_m3", variable),
    
    pollutant = variable,
  ) %>%
  variable_relabel(var = "pollutant") 
 
stop_no <- stops0 %>%
  distinct(location) %>% 
  mutate(location_no = row_number()) 

stops <- stops0 %>%
  mutate(
    #use doe neph (bscat x10^4 values) and pm2.5 calibration curve to calc pm2.5 from our neph readings
    value = ifelse(grepl("neph", variable), value*10^4*doe_slope + doe_int, value),
    variable = ifelse(grepl("neph", variable), "pm2.5_ug_m3", variable),
    
    day = wday(time,label = T, week_start = 1),
    hour = hour(time),
    pollutant = variable,
  ) %>%
  variable_relabel(var = "pollutant") %>%
  add_season(.date_var = "date") %>%
  
  left_join(stop_no)

  
```



# Drive Days

```{r}
#total 
stops %>%
  distinct(runname) %>%
  summarize(total_drive_days = n()) %>%
  kable(caption = "total drive days") %>%
  kable_styling()

#by instrument 
stops %>%
  distinct(runname, pollutant, ufp_range_nm) %>%
  group_by(pollutant, ufp_range_nm) %>%
  summarize(
    total_drive_days = n()
  ) %>%
  kable(caption = "total drive days by instrument") %>%
  kable_styling() 


```

```{r}
stops %>%
  mutate(
    hour = ifelse(hour==0, 24, hour)
  ) %>% 
  summarize(
    Earliest_sampling_hour = min(hour),
    Latest_sampling_hour = max(hour)
  ) %>%
  kable(caption = "earliest and latest sampling hour in campaign") %>%
  kable_styling()

```


# Stop medians

## Sample size

* this is for the stops the mobile platform completed. some instruments may have fewer readings (e.g., due to instrument issues)

```{r}
stops %>%
  distinct(runname, location) %>%  
  #visits per location
  group_by(location) %>%
  summarize(
    N_visits = n()
  ) %>%   
  #disgtribution of number of site visits
  summarize(
    N = n(),
    Min = min(N_visits),
    Q25 = quantile(N_visits, 0.25),
    Median = quantile(N_visits, 0.50),
    Q75 = quantile(N_visits, 0.95),
    Max = max(N_visits),
  ) %>%
  kable(caption = "number of visits to each location") %>%
  kable_styling() 
  
```

number of visits at different times of the day    
* showing example for UFPs since all pollutant trend were similar    
* sites were visited every season and day of the week    
* sites were more likely to be visited during certain hours: early morning, early afternoon, and evening hours


```{r}

print("total number of site visits")

stops %>%
  #show UFPs as an example
  filter(variable == "pnc_noscreen") %>%

  gather(temporal_var, temporal_value, season, day, hour) %>%
  mutate(
    temporal_var = factor(temporal_var, levels = c("season", "day", "hour")),
    temporal_value = factor(temporal_value, 
                            levels = c(levels(stops$season), levels(stops$day), c(0:23) )
                              )
    ) %>%  
  # group_by(temporal_var, temporal_value, location, location_no, variable) %>%
  # summarize(count = n()) %>%
  
  ungroup() %>%
  mutate(location_no = factor(location_no)) %>%
  
  ggplot(aes(x=temporal_value,
             y=location_no,
             )) + 
  facet_grid(pollutant~temporal_var, scales="free") + 
  geom_bin2d(aes(),
             #1 group per unit in the x and y directions
             #bins=1
             ) + 
  
  #scale_y_discrete(breaks = c(1, seq(50, 300, 50))) +
  
  labs(
       x = "Sampling Time" ,
       y="Site",
       fill = "No. Visits"
       )

```


## Temporal Trends 

notes   
* note, plotting non-extreme values, otherwise can't see any trends in the plots    
  - if on log scale, values <=0 are dropped (e.g., many for BC since aethalometer instrument is noisy)   
* note that different sites were visited at different times, so there could be some spatial confounding   
  
* **boxplots whiskers are for quantiles 2.5 and 97.5** 

### --> NanoScan (10-420 nm) has higher readings than discmini (10-700 nm)?

by month 


```{r, fig.height=12}
# by month

print("Median stop concentrations over time")

stops %>%
    mutate(month = factor(month(time, label = T))) %>%
    
    #calc stats for alternative boxplots
    group_by(pollutant, ufp_range_nm, month) %>%
    alt_boxplot(var = "value") %>%  #View()
    
    ggplot(aes(x=month, col=ufp_range_nm)) +
    geom_boxplot(aes(ymin=Qmin, lower=Q25, middle=Q50, upper=Q75, ymax=Qmax), stat = "identity") + 

    # # what is group=1?
    stat_summary(aes(y=Q50, group=ufp_range_nm), fun="mean", geom="line",  
                 size=1
                 ) +
    
  facet_wrap(~pollutant, scales="free") +
    labs(y= "Value",
         x="Month",
         col = "UFP Range (nm)"
         )  

```

by weekday

```{r, fig.height=12}

stops %>%
    #calc stats for alternative boxplots
    group_by(pollutant, ufp_range_nm, day) %>%
    alt_boxplot(var = "value") %>%  
    
    ggplot(aes(x=day, col=ufp_range_nm)) +
    geom_boxplot(aes(ymin=Qmin, lower=Q25, middle=Q50, upper=Q75, ymax=Qmax), stat = "identity") + 

    # what is group=1?
    stat_summary(aes(y=Q50, group=ufp_range_nm,), fun="mean", geom="line", #col="blue"
                 size=1
                 ) +
    
    facet_wrap(~pollutant, scales="free") +
    
    labs(y= "Value",
         x="Day",
         col = "UFP Range (nm)"
    )
                   
```

by hour 

```{r}

stops %>%
    #calc stats for alternative boxplots
    group_by(pollutant, ufp_range_nm, hour) %>%
    alt_boxplot(var = "value") %>%  
    
    ggplot(aes(x=factor(hour), col=ufp_range_nm)) +
    geom_boxplot(aes(ymin=Qmin, lower=Q25, middle=Q50, upper=Q75, ymax=Qmax), stat = "identity") + 

    # # what is group=1?
     stat_summary(aes(y=Q50, group=ufp_range_nm,), fun="mean", geom="line",  #col="blue"
                  size=1
                  ) +
    
  facet_wrap(~pollutant, scales="free") +
    
    labs(y= "Value",
         x= "Hour",
         col = "UFP Range (nm)"
         )  

```

 
## Spatial trends

* example using 10 sites per high, medium and low variability site, based on that site's stop concentration SD (based on tertile quantile break points: 0.33, 0.66)

within site variability 

* note that a few (7) BC measuresments are dropped if on log10 x axis scale b/c these are 0 ng/m3.

```{r, fig.height=10}

print("Distribution of median stop concentrations per site. Each colored line is a site with approximately 29 concentrations.")

stops %>%
  ggplot(aes(x=value, col=location)) + 
  facet_wrap(~pollutant+ufp_range_nm, scales="free") + 
  #use this instead of geom_density b/c can change line transparancy
  geom_line(stat="density", alpha=0.2) +
  scale_x_log10() +
  guides(col = F) 

```


```{r, fig.height=18, eval=F}
#hard to see any patterns b/c site are not organized.
#sample a few sites w/ low, medium and high variability
loc_sample <- stops %>%
  group_by(location) %>%
  mutate(sd = sd(value)) %>%
  ungroup() %>%
  mutate(
    sd_lvl = ifelse(sd >= quantile(sd, 0.66), "high",
                      ifelse(sd < quantile(sd, 0.66) & sd >= quantile(sd, 0.33), "medium",
                                                               "low")
                      )
    ) %>% 
  #one row per site
  distinct(location, sd_lvl) %>%
  #take random sample from each SD level
  group_by(sd_lvl) %>%
  sample_n(10)  
    
 stops %>%
   #only keep sample of sites
   right_join(loc_sample) %>%  
   
  #calc stats for alternative boxplots
  group_by(pollutant, ufp_range_nm, location) %>%
  alt_boxplot(var = "value") %>%  
  
  ggplot(aes(y=factor(location), col=ufp_range_nm)) +
  geom_boxplot(aes(xmin=Qmin, xlower=Q25, xmiddle=Q50, xupper=Q75, xmax=Qmax), stat = "identity") +
   facet_wrap(~pollutant, scales="free") +
    
    labs(y= "Site",
         x= "Value"
         )  


```

 
# Raw data

### --> look at this? note that data are not cleaned for use at the second level

* note that ~ 24k rows of data are removed for plots on log10 x axes b/c values <=0 

```{r}
sec %>%
  ggplot(aes(x=value, col=location)) + 
  facet_wrap(~pollutant+ufp_range_nm, scales="free") + 
  #use this instead of geom_density b/c can change line transparancy
  geom_line(stat="density", alpha=0.2) +
  scale_x_log10() +
  guides(col = F) 

```


boxplots with whiskers at 5 and 95

```{r, fig.height=10, eval=F}
print("Distribution of raw second readings for a set of example sites (same as above)")

sec %>%
  variable_relabel(var = "variable") %>%
  
 #only keep sample of sites
 right_join(loc_sample) %>% 
  
  #calc stats for alternative boxplots
  group_by(variable, location) %>%
  alt_boxplot(var = "value") %>%   
  
  ggplot(aes(y=location)) +
  geom_boxplot(aes(xmin=Qmin, xlower=Q25, xmiddle=Q50, xupper=Q75, xmax=Qmax), stat = "identity") +
  facet_wrap(~variable, scales="free") +
  
  labs(y="Location",
       x="Value"
       )
 


```

### --> ? what to do with this? 

```{r}
#IQR table 

sec %>%
  #calc stats for alternative boxplots
  group_by(pollutant, ufp_range_nm, location) %>%
  summarize(
    IQR = IQR(value)
  ) %>%  #View()
  summarize(
    Min = min(IQR),
    Median = median(IQR),
    Max = max(IQR)
  ) %>%
  mutate_if(is.numeric, 
            #if values have decimal values, only show 3 sigfigs, though not necessarily in scientific notation
            ~format(., digits = 3, #scientific = T,
                    ) 
            #~ signif(., 3) #~ as.character( signif(., 3) )
            ) %>%
  kable(caption = "Distribution of IQR values for the second data", 
        
        #digits=2
        ) %>%
  kable_styling()
  
  
```


# Annual averages


* calculating a simple mean annual average using all the available site data 

-summary tables/plots 

```{r}
annual <- stops %>%
  group_by(variable, pollutant, ufp_range_nm, location, site) %>%
  summarize(value = mean(value)) %>%
  ungroup() %>%
  #add location lat/long
  left_join(locations)


```


```{r}
#table of distribution
annual %>%
  mutate(
    pollutant = as.character(pollutant),
    pollutant = ifelse(grepl("UFP", pollutant), paste0(pollutant, ", ", ufp_range_nm), pollutant)
         ) %>%
  group_by(Pollutant=pollutant) %>%
  summarize(
    N = n(),
    Min = min(value),
    Q25 = quantile(value, 0.25),
    Median = quantile(value, 0.50),
    Q75 = quantile(value, 0.75),
    Max = max(value),
  ) %>%
  mutate_if(is.numeric,  ~format(., digits = 3, )) %>%
  kable(caption = "Annual average pollutant concentrations at monitoring sites. N = number of sites.") %>%
  kable_styling()
  
```

**these are traditional boxplots (hence why you see outliers)**

```{r, fig.height=10}
#boxplots
annual %>%
  ggplot(aes(x=pollutant, y=value, col=ufp_range_nm)) + 
  facet_wrap(~pollutant, scales= "free") + 
  geom_boxplot() + 
  
  labs(x="Pollutant",
       y = "Value",
       col = "UFP Range (nm)"
       )

```

## Maps 

```{r}
 
annual2 <- st_as_sf(annual, coords = c("longitude", "latitude"),  remove = F, crs=crs_deg) #%>% st_transform(crs = crs_m)

#need bbox w/ lat/long coordinates
bbox <- st_bbox(st_transform(st_buffer(st_transform(monitoring_area_shp, crs_m), 10000), crs_deg))

names(bbox) <- c("left", "bottom", "right", "top")

map0 <- suppressMessages(get_stamenmap(
  bbox = bbox,
  #zoom = 7,
 maptype = "terrain", 
  ))

# map0 %>% ggmap()

unique_p <- distinct(annual[c("pollutant", "ufp_range_nm")])

non_ufp_p <- str_subset(string =  unique_p$pollutant, 
           pattern = "UFP", 
           negate = T)
ufp_p <- str_subset(string =  unique_p$pollutant, 
           pattern = "UFP", 
           negate = F) %>% unique()



```

```{r, fig.height=18}
#non-UFPs. each pollutant has its own legend

p <- list()

for(i in seq_along(non_ufp_p)) {
   #i=1
   
  p[[i]] <- ggmap(ggmap = map0) +
    #monitoring area
    geom_sf(data=monitoring_area_shp, inherit.aes = F, aes(fill = "Monitoring Area"),
            alpha=0.1,
            #reduce/eliminate outline
            lwd = 0.1,
            ) +
    
    #annual averages
    geom_point(data=filter(annual, 
                         #pollutant == unique_p$pollutant[i] & ufp_range_nm ==  unique_p$ufp_range_nm[i]), 
                         pollutant == non_ufp_p[i]),
             aes(x=longitude, y=latitude, col=value), 
             inherit.aes = F)  +
    scale_color_gradient(name = "Conc",
                        low = "yellow", high = "red",
                        ) +
    facet_wrap(~pollutant+ufp_range_nm)  + 
    
    # add scale & N arrow to top rught
    annotation_scale(location = "tr") +
    annotation_north_arrow(location = "tr",
                           #point towards North Pole
                           which_north = "true",
                           pad_y = unit(0.5, "in"),
                           style = north_arrow_fancy_orienteering
                           ) +
  theme_bw() +
  theme(
    legend.justification=c(0,1),  
    legend.position=c(0,1),  
    legend.background =  element_blank()
    ) +
    #arrange legend order
  guides(col = guide_colorbar(order = 2),
              fill = guide_legend(order = 1)
         ) +
  labs(x = "Longitude",
       y = "Latitude",
       fill = ""
       ) 
  
  p[[i]]
  
  }

ggarrange(plotlist = p, ncol = 2, nrow = 2)

```

* UFP predictions seem to vary substantially depending on what instrument was used?    
  * note, people wouldn't normally use screened PTRAKS (50-1,000 nm particles) to measure "UFPs"

```{r, fig.height=18}
#UFPs. these share the same legend

ggmap(ggmap = map0) +
    #monitoring area
    geom_sf(data=monitoring_area_shp, inherit.aes = F, aes(fill = "Monitoring Area"),
            alpha=0.1,
            #reduce/eliminate outline
            lwd = 0.1,
            ) +
    
    #annual averages
    geom_point(data=filter(annual, pollutant == ufp_p, 
                           #if we don't included screened PTRAKs
                           #variable != "pnc_screen"
                           
                           ),
             aes(x=longitude, y=latitude, col=value), 
             inherit.aes = F)  +
  scale_color_gradient(name = "Conc",
                       low = "yellow", high = "red",
                       ) +
    facet_wrap(~pollutant+ufp_range_nm, nrow = 2)  + 
    
    # add scale & N arrow to top rught
    annotation_scale(location = "tr") +
    annotation_north_arrow(location = "tr",
                           #point towards North Pole
                           which_north = "true",
                           pad_y = unit(0.5, "in"),
                           style = north_arrow_fancy_orienteering
                           ) +
  theme_bw() +
  theme(
    legend.justification=c(0,1),  
    legend.position=c(0,1),  
    legend.background =  element_blank()
    ) +
    #arrange legend order
  guides(col = guide_colorbar(order = 2),
              fill = guide_legend(order = 1)
         ) +
  labs(x = "Longitude",
       y = "Latitude",
       fill = ""
       ) 

```

