---
title: "Campaign Data Summary"
author: "Magali Blanco"
date: ' `r Sys.time()` '
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
    number_sections: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
editor_options: 
  chunk_output_type: console
---

```{r, echo=F}
##merge with github. always do these 4 steps all together. enter the following into the Terminal.
# 
###a. git pull origin master #do this before start editing & before you're read to upload changes again
###b. git add _____.Rmd 
###c. git commit -m "___my commit message in parentheses" 
###d. git push origin master
# 
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      cache=F, cache.comments = F, 
                      message = F, warning = F, 
                      tidy.opts=list(width.cutoff=60), tidy=TRUE,
                      #fig.height = 5, fig.width = 8
                      fig.height = 6, fig.width = 10
                      )  

# # Clear workspace of all objects and unload all extra (non-base) packages
# rm(list = ls(all = TRUE))
# if (!is.null(sessionInfo()$otherPkgs)) {
#   res <- suppressWarnings(
#     lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
#       detach, character.only=TRUE, unload=TRUE, force=TRUE))
# }

pacman::p_load(knitr, kableExtra, 
               ggpubr, tidyverse,
               ggrepel, #geom_label_repel
               ggmap, sf, ggspatial, #mapping...adding scales, N arrows
               ggrepel, #avoid overlapping labels
               units, #convert between e.g., m to km
               #time series data 
               lubridate,
               
               fmsb #, #percentile()
               #GGally #ggpairs()
               )    
 

set.seed(1)

image_path <- file.path("..", "Manuscript", "Images")
theme_set(theme(legend.position = "bottom"))

```


# Upload data

common variables

```{r}
low_boxplot_q <- 0.05
high_boxplot_q <- 0.95

```


###--> add uncalibrated data/other data for sensitivity analyses 

```{r}
#functions
source("0_Functions.R")

#only need location table
load(file.path("Data", "Original", "locations_table.rda"))

#definitions: ap 
load(file.path("Data", "Output", "common_vars.rda"))

#stop data
sec0 <- readRDS(file.path("Data", "Output", "second_data.rda"))
stops0 <- readRDS(file.path("Data", "Output", "stop_data.rda"))

```

```{r}
#mapping
crs_m <- 32148
crs_deg <- 4326 #WGS84. in decimal degrees

#shapefiles
monitoring_area_shp <- read_sf(file.path("..", "GIS", "Shapefiles", "Monitoring", "monitoring_area_filled_in.shp")) %>%
  #convert from 4269 
  st_transform(crs_deg)

```


```{r}
# sec <- sec0 %>%
#   mutate(
#     
#     #use doe neph (bscat x10^4 values) and pm2.5 calibration curve to calc pm2.5 from our neph readings
#     value = ifelse(grepl("neph", variable), value*10^4*slope + int, value),
#     variable = ifelse(grepl("neph", variable), "pm2.5_ug_m3", variable),
#     
#     pollutant = variable,
#   ) %>%
#   variable_relabel(var = "pollutant") 
 
keep_names <- setdiff(names(stops0), c("instrument_id", "primary_instrument", "ufp_range_nm")) #"value_c1", "value_raw", 

# calculate UFPs 20-50 nm using PTRAK data
ufp_20_50 <- stops0 %>%
  filter(grepl("PMPT", instrument_id)) %>%
  select(keep_names) %>%
  gather("value_v", "value", contains("value")) %>%
  mutate(variable = paste0(variable, "_", value_v)) %>%
  select(-value_v) %>%
  spread(variable, value) %>%
  drop_na() %>%
  mutate(
    pnc_20_50_value = pnc_noscreen_value - pnc_screen_value,
    pnc_20_50_value_c1 = pnc_noscreen_value_c1 - pnc_screen_value_c1,
    pnc_20_50_value_raw = pnc_noscreen_value_raw - pnc_screen_value_raw
  ) %>%
  gather("variable", "value", contains("pnc_")) %>%
  mutate(
    value_v = str_extract(variable, "v.*"),
    variable = sub("_v.*", "", variable )
  ) %>%
  spread(value_v, value) %>%
  filter(grepl("20_50", variable))

# add stop #s
stop_no <- stops0 %>%
  distinct(location) %>% 
  mutate(location_no = row_number()) 

stops <- stops0 %>%
  select(keep_names) %>%
  #add new UFP estimates
  rbind(ufp_20_50) %>%
  # add stop #
  left_join(stop_no) %>%
  # add other labels/temporal variables
  mutate(
    variable = ifelse(grepl("neph", variable), "pm2.5_ug_m3", variable),
    day = wday(time,label = T, week_start = 1),
    hour = hour(time),
    pollutant = variable,
  ) %>%
  variable_relabel(var = "pollutant") %>%
  add_season(.date_var = "date") 
  
```


# Drive Days

```{r}
#total 
stops %>%
  distinct(runname) %>%  
  summarize(total_drive_days = n()) %>%
  kable(caption = "total drive days") %>%
  kable_styling()

## by instrument
stops %>%
  group_by(pollutant, ufp_range_nm) %>%
  distinct(runname) %>%  
  summarize(total_drive_days = n()) %>%
  kable(caption = "total drive days, by instrument") %>%
  kable_styling()

```

```{r}
stops %>%
  mutate(hour = ifelse(hour==0, 24, hour)) %>% 
  summarize(
    Earliest_sampling_hour = min(hour),
    Latest_sampling_hour = max(hour)
  ) %>%
  kable(caption = "earliest and latest sampling hour in campaign") %>%
  kable_styling()

```


# Stop medians

## Sample size

* this is for the stops the mobile platform completed. some instruments may have fewer readings (e.g., due to instrument issues)

```{r}
stops %>%
  distinct(runname, location) %>%  
  #visits per location
  group_by(location) %>%
  summarize(
    N_visits = n()
  ) %>%   
  #disgtribution of number of site visits
  summarize(
    N = n(),
    Min = min(N_visits),
    Q25 = quantile(N_visits, 0.25),
    Mean = mean(N_visits),
    SD = sd(N_visits),
    Median = quantile(N_visits, 0.50),
    IQR = IQR(N_visits),
    Q75 = quantile(N_visits, 0.95),
    Max = max(N_visits),
  ) %>%
  kable(caption = "number of visits to each location", 
        digits = 1
        ) %>%
  kable_styling() 
  
```

number of visits at different times of the day    
* showing example for UFPs since all pollutant trend were similar    
* sites were visited every season and day of the week    
* sites were more likely to be visited during certain hours: early morning, early afternoon, and evening hours


```{r}

print("total number of site visits")

stops %>%
  #show UFPs as an example
  filter(variable == "pnc_noscreen") %>%

  gather(temporal_var, temporal_value, season, day, hour) %>%
  mutate(
    temporal_var = factor(temporal_var, levels = c("season", "day", "hour")),
    temporal_value = factor(temporal_value, 
                            levels = c(levels(stops$season), levels(stops$day), c(0:23) )
                              )
    ) %>%  
  # group_by(temporal_var, temporal_value, location, location_no, variable) %>%
  # summarize(count = n()) %>%
  
  ungroup() %>%
  mutate(location_no = factor(location_no)) %>%
  
  ggplot(aes(x=temporal_value,
             y=location_no,
             )) + 
  facet_grid(pollutant~temporal_var, scales="free") + 
  geom_bin2d(aes(),
             #1 group per unit in the x and y directions
             #bins=1
             ) + 
  
  #scale_y_discrete(breaks = c(1, seq(50, 300, 50))) +
  
  labs(
       x = "Sampling Time" ,
       y="Site",
       fill = "No. Visits"
       )

```


## Spatial trends



```{r, eval=F, fig.height=10}
# * example using 10 sites per high, medium and low variability site, based on that site's stop concentration SD (based on tertile quantile break points: 0.33, 0.66)
# 
# * note that a few (7) BC measuresments are dropped if on log10 x axis scale b/c these are 0 ng/m3.

print("Distribution of median stop concentrations per site. Each colored line is a site with approximately 29 concentrations.")

stops %>%
  ggplot(aes(x=value, col=location)) + 
  facet_wrap(~pollutant+ufp_range_nm, scales="free") + 
  #use this instead of geom_density b/c can change line transparancy
  geom_line(stat="density", alpha=0.2) +
  scale_x_log10() +
  guides(col = F) 

```

* boxplots of site visit medians.    
* whiskers are for the quantiles: `r low_boxplot_q` and `r high_boxplot_q`

```{r}
unique_pollutants <- unique(stops$pollutant)

location_quant <- stops %>%
  #calc stats for alternative boxplots
  group_by(pollutant, ufp_range_nm, location, location_no) %>%
  alt_boxplot(var = "value", min_q = low_boxplot_q, max_q = high_boxplot_q)

```

```{r, eval=F}

for (i in seq_along(unique_pollutants)) {
  #i=1
    p <- location_quant %>%
      filter(pollutant == unique_pollutants[i]) %>%
    
      ggplot(aes(y=factor(location_no), col=ufp_range_nm)) +
      geom_boxplot(aes(xmin=Qmin, xlower=Q25, xmiddle=Q50, xupper=Q75, xmax=Qmax), stat = "identity") +
       facet_wrap(~pollutant+ufp_range_nm, nrow = 1) +
      scale_y_discrete(breaks = seq(10,310, 10)) +
      labs(y= "Site No.",
             x= "Value"
             )  +
      theme(legend.position = "none")
    
    print(p)
    
    ggsave(p, file = file.path(image_path, "SI", paste0("visit_boxplots_", sub(" .*", "", unique_pollutants[i]), ".png")),
         height = 11, width = 9
         )
}


```

```{r}
#non-ufp plots
location_quant %>%
  filter(!grepl("UFP", pollutant)) %>%
  ggplot(aes(y=factor(location_no), col=ufp_range_nm)) +
  geom_boxplot(aes(xmin=Qmin, xlower=Q25, xmiddle=Q50, xupper=Q75, xmax=Qmax), stat = "identity") +
  facet_wrap(~pollutant+ufp_range_nm, scales="free", nrow = 1) +
  scale_y_discrete(breaks = seq(10,310, 10)) +
  labs(y= "Site No.",
         x= "Value")  +
  theme(legend.position = "none") #+ scale_x_log10()
    
ggsave(file = file.path(image_path, "SI", "visit_boxplots_non_ufp.png"),
         height = 9, width = 7)

#ufp plots
location_quant %>%
  filter(grepl("UFP", pollutant),
         #!grepl("50-1,000", ufp_range_nm)
         ) %>%
  ggplot(aes(y=factor(location_no), col=ufp_range_nm)) +
  geom_boxplot(aes(xmin=Qmin, xlower=Q25, xmiddle=Q50, xupper=Q75, xmax=Qmax), stat = "identity") +
  facet_wrap(~pollutant+ufp_range_nm, #scales="free", 
             nrow = 1) +
  scale_y_discrete(breaks = seq(10, 310, 10)) +
  theme(legend.position = "none") + 
  labs(y= "Site No.",
         x= "Value") + #geom_vline(xintercept = 20e3, linetype=3, alpha=0.5)
  #theme(axis.text.x = element_text(size = 11))
    scale_x_continuous(guide = guide_axis(check.overlap = TRUE,
                                          #n.dodge=2
                                          ))
  
ggsave(file = file.path(image_path, "SI", "visit_boxplots_ufp.png"),
         height = 9, width = 7)

```



```{r, fig.height=18, eval=F}
#hard to see any patterns b/c site are not organized.
#sample a few sites w/ low, medium and high variability
loc_sample <- stops %>%
  group_by(location) %>%
  mutate(sd = sd(value)) %>%
  ungroup() %>%
  mutate(
    sd_lvl = ifelse(sd >= quantile(sd, 0.66), "high",
                      ifelse(sd < quantile(sd, 0.66) & sd >= quantile(sd, 0.33), "medium",
                                                               "low")
                      )
    ) %>% 
  #one row per site
  distinct(location, sd_lvl) %>%
  #take random sample from each SD level
  group_by(sd_lvl) %>%
  sample_n(10)  
    
 stops %>%
   #only keep sample of sites
   right_join(loc_sample) %>%  
   #calc stats for alternative boxplots
  group_by(pollutant, ufp_range_nm, location) %>%
  alt_boxplot(var = "value") %>%  

  ggplot(aes(y=factor(location), col=ufp_range_nm)) +
  geom_boxplot(aes(xmin=Qmin, xlower=Q25, xmiddle=Q50, xupper=Q75, xmax=Qmax), stat = "identity") +
   facet_wrap(~pollutant, scales="free") +
    labs(y= "Site",
         x= "Value"
         )  


```

## Temporal Trends 

notes   
* note, plotting non-extreme values, otherwise can't see any trends in the plots    
  - if on log scale, values <=0 are dropped (e.g., many for BC since aethalometer instrument is noisy)   
* note that different sites were visited at different times, so there could be some spatial confounding   
  

* NanoScan (10-420 nm) has higher readings than discmini (10-700 nm) b/c it is more efficient at measuring the smallest particles due to the difference in technologies. The Nanoscan measures optical diameter after alcohol condensation to grow the particles, whereas the disk-mini measures electrical mobility diameter based on the current produced when the particles are captured on a oppositely charged surface.  The electrostatic method has lower detection efficiencies for these very small particles than does the condensation/optical method.

by month 


```{r, fig.height=12}
# by month

print("Median site visit concentrations by season. Lines connect median monthly concentrations.")

stops %>%
    mutate(month =month(time, label = T),
           ) %>%  
    #calc stats for alternative boxplots
    group_by(pollutant, ufp_range_nm, month, season) %>%
    alt_boxplot(var = "value", min_q = low_boxplot_q, max_q = high_boxplot_q) %>% 
    
    ggplot(aes(x=month, col=ufp_range_nm)) +
    geom_boxplot(aes(ymin=Qmin, lower=Q25, middle=Q50, upper=Q75, ymax=Qmax), stat = "identity") + 
    stat_summary(aes(y=Q50, group=ufp_range_nm), fun="mean", geom="line", size=1) +
  facet_grid(pollutant~season, scales="free") +
    labs(y= "Value",
         x="Month",
         col = "UFP Range (nm)"
         )  

ggsave(file = file.path(image_path, "SI", "temporal_season.png"),
         height = 9, width = 7)

```

by weekday

```{r, fig.height=12}
print("Median site visit concentrations by day of the week. Lines connect median monthly concentrations.")

stops %>%
  mutate(dow = ifelse(day %in% c("Sat", "Sun"), "Weekend", "Weekday")) %>%
  #calc stats for alternative boxplots
  group_by(pollutant, ufp_range_nm, day, dow) %>%
  alt_boxplot(var = "value") %>%  
  
  ggplot(aes(x=day, col=ufp_range_nm)) +
  geom_boxplot(aes(ymin=Qmin, lower=Q25, middle=Q50, upper=Q75, ymax=Qmax), stat = "identity") + 
  stat_summary(aes(y=Q50, group=ufp_range_nm,), fun="mean", geom="line", size=1) +
  facet_grid(pollutant~dow, scales="free", space = "free_x") +
  labs(y= "Value",
       x="Day",
       col = "UFP Range (nm)"
  )


ggsave(file = file.path(image_path, "SI", "temporal_day.png"),
         height = 9, width = 7)                 
```

by hour 

```{r}

stops %>%
  mutate(hour = factor(hour, levels = c(4:23, 0))) %>%
    #calc stats for alternative boxplots
    group_by(pollutant, ufp_range_nm, hour) %>%
    alt_boxplot(var = "value") %>%  
    
    ggplot(aes(x=hour, col=ufp_range_nm)) +
    geom_boxplot(aes(ymin=Qmin, lower=Q25, middle=Q50, upper=Q75, ymax=Qmax), stat = "identity") + 
     stat_summary(aes(y=Q50, group=ufp_range_nm,), fun="mean", geom="line",size=1) +
    
  facet_wrap(~pollutant, scales="free", nrow = 3) +
    
    labs(y= "Value",
         x= "Hour",
         col = "UFP Range (nm)"
         )  

ggsave(file = file.path(image_path, "SI", "temporal_hour.png"),
         height = 9, width = 7)                 

```

 

 
# Raw data

* note that ~ 24k rows of data are removed for plots on log10 x axes b/c values <=0 

```{r,  fig.height=10}
# sec %>%
#   ggplot(aes(x=value, col=location)) + 
#   facet_wrap(~pollutant+ufp_range_nm, scales="free") + 
#   #use this instead of geom_density b/c can change line transparancy
#   geom_line(stat="density", alpha=0.2) +
#   scale_x_log10() +
#   guides(col = F) 

```


boxplots with whiskers at 5 and 95

```{r, fig.height=10, eval=F}
# print("Distribution of raw second readings for a set of example sites (same as above)")
# 
# sec %>%
#   variable_relabel(var = "variable") %>%
#   
#  #only keep sample of sites
#  right_join(loc_sample) %>% 
#   
#   #calc stats for alternative boxplots
#   group_by(variable, location) %>%
#   alt_boxplot(var = "value") %>%   
#   
#   ggplot(aes(y=location)) +
#   geom_boxplot(aes(xmin=Qmin, xlower=Q25, xmiddle=Q50, xupper=Q75, xmax=Qmax), stat = "identity") +
#   facet_wrap(~variable, scales="free") +
#   
#   labs(y="Location",
#        x="Value"
#        )
 


```


```{r}
# ## --> ? what to do with this? 
# 
# #IQR table 
# 
# sec %>%
#   #calc stats for alternative boxplots
#   group_by(pollutant, ufp_range_nm, location) %>%
#   summarize(
#     IQR = IQR(value)
#   ) %>%  #View()
#   summarize(
#     Min = min(IQR),
#     Median = median(IQR),
#     Max = max(IQR)
#   ) %>%
#   mutate_if(is.numeric, 
#             #if values have decimal values, only show 3 sigfigs, though not necessarily in scientific notation
#             ~format(., digits = 3, #scientific = T,
#                     ) 
  #           #~ signif(., 3) #~ as.character( signif(., 3) )
  #           ) %>%
  # kable(caption = "Distribution of IQR values for the second data", 
  #       
  #       #digits=2
  #       ) %>%
  # kable_styling()
  
  
```


# Annual averages

* calculating a simple mean annual average using all the available site data 

### --> also do this w/ uncalibrated data (? & zero data?) & compare annual averages  


## Tables  

```{r}
annual <- stops %>%
  group_by(variable, pollutant, ufp_range_nm, location) %>%
  summarize(value = mean(value)) %>%
  ungroup() %>%
  #add location lat/long
  left_join(locations)


```


```{r}
#table of distribution
annual %>%
  mutate(
    pollutant = as.character(pollutant),
    pollutant = ifelse(grepl("UFP", pollutant), paste0(pollutant, ", ", ufp_range_nm), pollutant)
         ) %>%
  group_by(Pollutant=pollutant) %>%
  summarize(
    #N = n(),
    Min = min(value),
    Q25 = quantile(value, 0.25),
    Median = quantile(value, 0.50),
    Q75 = quantile(value, 0.75),
    Max = max(value),
  ) %>%
  mutate_if(is.numeric,  ~format(., digits = 0, scientific = F, big.mark = ",")) %>%
  kable(caption = "Distribution of annual average pollutant concentrations at monitoring sites (N = 309)") %>%
  kable_styling()
  
```

**these are traditional boxplots (hence why you see outliers)**

```{r, fig.height=10}
print("Annual average site concentrations")

#boxplots
annual %>%
  ggplot(aes(x=pollutant, y=value, col=ufp_range_nm)) + 
  facet_wrap(~pollutant, scales= "free") + 
  geom_boxplot() + 
  
  labs(x="Pollutant",
       y = "Value",
       col = "UFP Range (nm)"
       )

ggsave(file.path(image_path, "annual_avg.png"), height = 6, width = 7)

```

## Maps 


```{r}
 
annual2 <- st_as_sf(annual, coords = c("longitude", "latitude"),  remove = F, crs=crs_deg) #%>% st_transform(crs = crs_m)

#need bbox w/ lat/long coordinates
bbox <- st_bbox(st_transform(st_buffer(st_transform(monitoring_area_shp, crs_m), 10000), crs_deg))

names(bbox) <- c("left", "bottom", "right", "top")

map0 <- suppressMessages(get_stamenmap(
  bbox = bbox,
  #zoom = 7,
 maptype = "terrain"
  ))

# map0 %>% ggmap()

unique_p <- distinct(annual[c("pollutant", "ufp_range_nm")])

non_ufp_p <- str_subset(string =  unique_p$pollutant, 
           pattern = "UFP", 
           negate = T)
ufp_p <- str_subset(string =  unique_p$pollutant, 
           pattern = "UFP", 
           negate = F) %>% unique()



```

### --> ? make background transparant? fix x-labels

```{r, fig.height=18}
#non-UFPs. each pollutant has its own legend

p <- list() 

for(i in seq_along(non_ufp_p)) {
   #i=1
   
  p[[i]] <- ggmap(ggmap = map0) +
    #monitoring area
    geom_sf(data=monitoring_area_shp, inherit.aes = F, aes(fill = "Monitoring Area"),
            alpha=0.1,
            #reduce/eliminate outline
            lwd = 0.1,
            ) +
    
    #annual averages
    geom_point(data=filter(annual,
                           pollutant == non_ufp_p[i]),
             aes(x=longitude, y=latitude, col=value), 
             inherit.aes = F, size=2)  +
    scale_color_gradient(name = "Conc",
                        low = "yellow", high = "red",
                        ) +
    facet_wrap(~pollutant)  + 
    # add scale & N arrow to top rught
    annotation_scale(location = "tr") +
    annotation_north_arrow(location = "tr",
                           #point towards North Pole
                           which_north = "true",
                           pad_y = unit(0.5, "in"),
                           style = north_arrow_fancy_orienteering
                           ) +
  theme_bw() +
  theme(
    legend.justification=c(0,1),  
    legend.position=c(0,1),  
    legend.background =  element_blank()
    ) +
    # scale_x_continuous(guide = guide_axis(check.overlap = TRUE)) + #n.dodge=2
    #scale_x_continuous(breaks = c(seq(-122.5, -121.9, 0.2))) +
    
    #arrange legend order
  guides(col = guide_colorbar(order = 2),
              fill = guide_legend(order = 1)
         ) +
  labs(x = "Longitude",
       y = "Latitude",
       fill = ""
       ) 
  
  p[[i]]
  
  }

ggarrange(plotlist = p, ncol = 2, nrow = 2)

ggsave(file.path(image_path, "map_non_ufp.png"),  height = 10, width = 8)

```

* UFP predictions seem to vary substantially depending on what instrument was used?    
  * note, people wouldn't normally use screened PTRAKS (50-1,000 nm particles) to measure "UFPs"

### --> plot UFPs as percentiles: percentile(). or standardized values?

```{r, fig.height=18}
#UFPs. these share the same legend

ggmap(ggmap = map0) +
    #monitoring area
    geom_sf(data=monitoring_area_shp, inherit.aes = F, aes(fill = "Monitoring Area"),
            alpha=0.1,
            #reduce/eliminate outline
            lwd = 0.1,
            ) +
    
    #annual averages
    geom_point(data=filter(annual, pollutant == ufp_p),
             aes(x=longitude, y=latitude, col=value), 
             inherit.aes = F)  +
  scale_color_gradient(name = "Conc",
                       low = "yellow", high = "red",
                       ) +
    facet_wrap(~pollutant+ufp_range_nm, nrow = 2)  + 
    
    # add scale & N arrow to top rught
    annotation_scale(location = "tr") +
    annotation_north_arrow(location = "tr",
                           #point towards North Pole
                           which_north = "true",
                           pad_y = unit(0.5, "in"),
                           style = north_arrow_fancy_orienteering
                           ) +
  theme_bw() +
  theme(
    legend.justification=c(0,1),  
    legend.position=c(0,1),  
    legend.background =  element_blank()
    ) +
    #arrange legend order
  guides(col = guide_colorbar(order = 2),
              fill = guide_legend(order = 1)
         ) +
  labs(x = "Longitude",
       y = "Latitude",
       fill = ""
       ) 


```


## Pollutant comparisons

### --> GGally::ggpairs() doesn't work? 

```{r, fig.height=10}

annual %>%
  select(variable, location, value) %>%  
  spread(variable, value) %>% 
  select(-location) %>%
  
  #GGally::ggpairs()
  plot(., )
  


```


* scatterplot comparison of PM2.5 vs other pollutant annual averages, using standardized values 

```{r, fig.height=12}
#note: could also use percentiles or regular scale

#make PM2.5 ref (x-axis) variable
y_vars <- setdiff(unique(annual$variable), "pm2.5_ug_m3")

annual %>%
  select(-c(pollutant, ufp_range_nm)) %>%
  
  #standardize values to plot them on the same scale
  group_by(variable) %>%
  mutate(
    #standardize values: (x-mean)/SD
    value = scale(value)
    # # alternative is to calculate percentiles     
    # value = percentile(value) 
         ) %>%  
  
  spread(variable, value) %>% 
  gather(variable, value, y_vars) %>% 
  #add cleaner labels
  variable_relabel(var = "variable") %>% # View()
  
  ggplot(aes(x=pm2.5_ug_m3, y=value, #col=variable, 
             col=ufp_range_nm, 
             linetype=ufp_range_nm #size = ufp_range_nm
             )) +
  
  geom_abline(slope = 1, intercept = 0, alpha=0.5) +
  geom_point(alpha=0.2) + 
  geom_smooth(se=F) +
  stat_cor(method = "pearson", 
           #only show R value, not p-val
           aes(label = ..rr.label..),
           #doesn't work??
           #r.accuracy = 0.01
           ) +
  
  #make x-y axis same
  facet_wrap_equal(~variable) +
  theme(aspect.ratio = 1) +
  
  labs(
    title = "Comparison of standardized annual averages at N=309 sites",
    y = "Other Pollutant Standardized Value",
    x = "PM2.5 Standardized Value",
    linetype = "UFP Size Range",
    col= "UFP Size Range",
    subtitle = "showing pearson R2" 
        )

```

* scatterplot comparison of different UFP instruments 

```{r}
#make NanoScan ref instrument
y_vars_ufp <- setdiff(unique(annual$variable[grepl("UFP", annual$pollutant)]), "ns_total_conc")

annual %>%
  #only look at UFPs
  filter(grepl("UFP", pollutant)) %>%
  
  select(-c(pollutant, ufp_range_nm)) %>%
  spread(variable, value) %>% 
  gather(variable, value, y_vars_ufp) %>% 
  #add cleaner labels
  variable_relabel(var = "variable") %>%  
  
  ggplot(aes(x=ns_total_conc, y=value, 
             col=ufp_range_nm, 
             )) +
  geom_abline(slope = 1, intercept = 0, alpha=0.5) +
  geom_point(alpha=0.2) + 
  geom_smooth(se=F) +
  stat_cor(method = "pearson", 
           #only show R value, not p-val
           aes(label = ..rr.label..),
           ) +
  #make x-y axis same
  facet_wrap_equal(~variable) +
  theme(aspect.ratio = 1) +
  
  labs(
    title = "Comparison of standardized annual averages at N=309 sites",
    y = "Other UFP Conc (pt/cm3)",
    x = "NanoScan (10-420 nm) Conc (pt/cm3)",
    col= "UFP Size Range",
    subtitle = "showing pearson R2" 
        )
```


### --> compare UFPs (x) to other instruments

```{r}

```




