---
title: "Campaign Data Summary"
author: "Magali Blanco"
date: ' `r Sys.time()` '
output:
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
    number_sections: true
    toc_float: true
    collapsed: false
    smooth_scroll: false
editor_options: 
  chunk_output_type: console
---

```{r, echo=F}
##merge with github. always do these 4 steps all together. enter the following into the Terminal.
# 
###a. git pull origin master #do this before start editing & before you're read to upload changes again
###b. git add _____.Rmd 
###c. git commit -m "___my commit message in parentheses" 
###d. git push origin master
# 
```


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      cache=F, cache.comments = F, 
                      message = F, warning = F, 
                      tidy.opts=list(width.cutoff=60), tidy=TRUE,
                      fig.height = 6, fig.width = 10
                      )  

# Clear workspace of all objects and unload all extra (non-base) packages
rm(list = ls(all = TRUE))
if (!is.null(sessionInfo()$otherPkgs)) {
  res <- suppressWarnings(
    lapply(paste('package:', names(sessionInfo()$otherPkgs), sep=""),
      detach, character.only=TRUE, unload=TRUE, force=TRUE))
}

pacman::p_load(knitr, kableExtra, broom,
               ggpubr, tidyverse,
               ggrepel, #geom_label_repel
               ggmap, sf, ggspatial, #mapping...adding scales, N arrows
               ggrepel, #avoid overlapping labels
               units, #convert between e.g., m to km
               #time series data 
               lubridate,
               
               fmsb, #percentile()
               GGally, #ggpairs()
               VCA, #anovaVCA()
               gstat, #variogram()
               ggpmisc, # for ggplot() geom="table"
               units, #convert between e.g., m to km
               colorspace #create color palates that "look good". e.g., rainbow_hcl(3)
               )    
 

set.seed(1)

theme_set(theme_bw())
theme_update(legend.position = "bottom")

image_path <- file.path("..", "Manuscript", "Images")

```


# Upload data

```{r}
#functions
source("0_Functions.R")

# 'locations' table
load(file.path("Data", "Original", "locations_table.rda"))

#definitions: ap 
load(file.path("Data", "Output", "common_vars.rda"))

#stop data
stops0 <- readRDS(file.path("Data", "Output", "stop_data.rda")) %>%
  select(-mean_value) %>% rename(value = median_value)


#geocovariates
#geo <- readRDS(file.path("~", "OneDrive - UW", "Documents","School","PhD", "Dissertation", "TRAP R Project","Data", "Aim 2", "Geocovariates", "cov_mm_preprocessed.rda"))

```

 
```{r}
#mapping
crs_m <- 32148
crs_deg <- 4326 #WGS84. in decimal degrees

#shapefiles
monitoring_area_shp <- readRDS(file.path("Data", "Output", "GIS", "monitoring_area_shp.rda")) %>%
  #convert from 4269 
  st_transform(crs_deg)

```

```{r}
#need bbox w/ lat/long coordinates
bbox <- st_bbox(st_transform(st_buffer(st_transform(monitoring_area_shp, crs_m), 10000), crs_deg))

names(bbox) <- c("left", "bottom", "right", "top")

map0 <- suppressMessages(get_stamenmap(
  bbox = bbox, 
  zoom = 11, 
  maptype = "toner-lite" #has airport symbol
  #maptype = "toner-background" #roads & water but no airport
    ))
#map0 %>% ggmap(darken = c(.4, "white"))


gis_path0 <- file.path("~", "OneDrive - UW", "Documents", "School", "PhD", "Dissertation", "GIS", "Shapefiles")
 
```

```{r}
 
keep_names <- setdiff(names(stops0), c("instrument_id", "primary_instrument", "ufp_range_nm")) #"value_c1", "value_raw", 

# add stop #s
stop_no <- stops0 %>%
  distinct(location) %>% 
  mutate(location_no = row_number()) 

stops0.1 <- stops0 %>%
  select(keep_names) %>%
   
  # add stop #
  left_join(stop_no) %>%
  # add other labels/temporal variables
  mutate(
    #variable = ifelse(grepl("neph", variable), "pm2.5_ug_m3", variable),
    day = wday(time,label = T, week_start = 1),
    hour = hour(time),
    pollutant = variable,
  ) %>%
  variable_relabel(var = "pollutant") %>%
  mutate(pollutant2 = ifelse(grepl("PNC", pollutant), 
                             paste0(pollutant, ", ", ufp_instrument),
                             as.character(pollutant)
                             )
         ) %>%
  add_season(.date_var = "date") 
  
```

```{r}

ptrak_bin_range <- c("20-36 nm", "36-1,000 nm")

```


```{r}
 stops <- stops0.1 %>%
  group_by(variable, location) %>%
  #winsorize median values  
  winsorize_fn(value = "value") %>%  
  ungroup() %>%
  select(-value) %>%
  #use winsorized values for all subsequent data description
  rename(value=win_value)
 
#save for future analyses with winsorized values
saveRDS(stops, file.path("Data", "Output", "stop_data_win_medians.rda"))

```


instrument names

```{r}
instruments <- stops %>%
  distinct(variable, pollutant, pollutant2) %>%
  mutate(
    Instrument = c("MA200", "Li-Cor", "M903", "DiSCmini", "Screened P-TRAK", "NanoScan", "CAPS", "P-TRAK")
  )

```



common variables

```{r}
low_boxplot_q <- 0.05
high_boxplot_q <- 0.95

# maps
## map labels
map_x_labels <- c(seq(-122.5, -121.9, 0.2)) #0.2
map_y_labels <- c(seq(47.2, 48, 0.2)) #%>% format(nsmall=1)
 
unique_pollutants <- unique(stops$pollutant) %>% sort()
unique_pollutants2 <- unique(stops$pollutant2) %>% sort()
uniqe_variables <-  
  relevel( 
  factor(unique(stops$variable)), 
                           ref = "no2" )

non_ufp_p <- str_subset(string =  unique_pollutants, 
           pattern = "PNC", 
           negate = T) %>%
  sort()


ufp_p <- str_subset(string =  unique_pollutants, 
           pattern = "PNC", 
           negate = F) %>% unique()


pollutant_levels <-  c("PNC (pt/cm3), P-TRAK", "PNC (pt/cm3), Screened P-TRAK", "PNC (pt/cm3), NanoScan", "PNC (pt/cm3), DiSCmini", "BC (ng/m3)", "NO2 (ppb)", "PM2.5 (ug/m3)",  "CO2 (ppm)")

pollutant_levels2 <-  c("PNC (pt/cm3)",  str_subset(string =  pollutant_levels, pattern = "PNC", negate = T))


save(pollutant_levels, pollutant_levels2, file = file.path("Data", "Output", "Objects", "pollutant_levels.rda"))


ufp_instruments <- c("P-TRAK", "Screened P-TRAK", "NanoScan", "DiSCmini")

```

# Route Statistics


```{r}
# annie's work 
on_road <- readRDS(file.path("Data", "Annie", "route_stats_08.02.21.rds"))

```



```{r}
# route distance
 
routes_shp <- read_sf(file.path(gis_path0, "Routes 190910", "All routes.shp")) %>%
  st_transform(crs_deg) %>%
  mutate(
    Route = as.factor(Route)
  )

```

 

```{r}
  #group_by(Route) %>%
#st_union(by_feature = T)
  #st_combine()

route_dist <- routes_shp %>%
  #calculate length of each piece of route string
  mutate(length_mi = as.numeric(set_units(st_length(.), "mi")),
         length_km = as.numeric(set_units(st_length(.), "km")),
         ) %>%
  st_drop_geometry() %>% as.data.frame() %>%
  #total route distnace
  group_by(Route) %>%
  summarise(length_mi = sum(length_mi),
            length_km = sum(length_km),
            )  
  

```



```{r}

#basic route stats
route_stats <- stops %>%
  mutate(
    Route = substr(runname, 14, nchar(runname))
  ) %>%
  group_by(Route) %>%
  summarize(
    stops = length(unique(location)),
    dates = length(unique(runname))
  ) %>%
  left_join(route_dist) %>%
  mutate(
    total_distance_mi = dates * length_mi,
    total_distance_km = dates * length_km
  ) 

#total row
tot <- route_stats %>%
  summarize(
    Route = "Total",
    stops = sum(stops),
    dates = max(dates),
    length_mi = sum(length_mi),
    length_km = sum(length_km),
    total_distance_mi = sum(total_distance_mi),
    total_distance_km = sum(total_distance_km)
    )

# combine
route_stats1 <- route_stats %>%
  rbind(tot) %>%
  left_join(on_road) %>%
  select(-no_road_segments)

route_stats1 %>%
  kable(caption = "Route statistics. Number of sampling dates includes makeup routes, which may have been slightly altered.", 
        col.names = c("Route", "No. Stops", "No. Sampling Dates", "Distance (mi)","Distance (km)", "Total Distance (mi)", "Total Distance (km)", "Median (IQR) Drive Time (hr)", "Total Drive Time (hr)" #names(on_road)[-1]
                      ),
        digits = 0, 
        format.args = list(big.mark = ",")
        ) %>%
  kable_styling()  


```


range estimates

```{r}
route_stats1 %>%
  slice(-10) %>%
  mutate_if(is.numeric, ~round(., 0)) %>%
  summarize(
    stops = paste0(min(stops), " - ", max(stops)),
    dates = paste0(min(dates), " - ", max(dates)),
    length_km = paste0(min(length_km), " - ", max(length_km)),
    length_mi = paste0(min(length_mi), " - ", max(length_mi)),
    total_distance_km = paste0(min(total_distance_km), " - ", max(total_distance_km)),
    total_distance_mi = paste0(min(total_distance_mi), " - ", max(total_distance_mi)),
    
  ) %>%
  kable(caption = "route-specific statistic ranges") %>%
  kable_styling()


```



# Drive Days

```{r}
#total 
stops %>%
  distinct(runname) %>%  
  summarize(total_drive_days = n()) %>%
  kable(caption = "total drive days") %>%
  kable_styling()

## by instrument
stops %>%
  group_by(pollutant, ufp_range_nm) %>%
  distinct(runname) %>%  
  summarize(total_drive_days = n()) %>%
  kable(caption = "total drive days, by instrument") %>%
  kable_styling()

```

```{r}
stops %>%
  mutate(hour = ifelse(hour==0, 24, hour)) %>% 
  summarize(
    Earliest_sampling_hour = min(hour),
    Latest_sampling_hour = max(hour)
  ) %>%
  kable(caption = "earliest and latest sampling hour in campaign") %>%
  kable_styling()

```


# Stop medians

## Sample size

* this is for the stops the mobile platform completed. some instruments may have fewer readings (e.g., due to instrument issues)

```{r}
stops %>%
  distinct(runname, location) %>%  
  #visits per location
  group_by(location) %>%
  summarize(
    N_visits = n()
  ) %>%   
  #disgtribution of number of site visits
  summarize(
    N = n(),
    Min = min(N_visits),
    Q25 = quantile(N_visits, 0.25),
    Mean = mean(N_visits),
    SD = sd(N_visits),
    Median = quantile(N_visits, 0.50),
    IQR = IQR(N_visits),
    Q75 = quantile(N_visits, 0.95),
    Max = max(N_visits),
  ) %>%
  kable(caption = "number of visits to each location", 
        digits = 1
        ) %>%
  kable_styling() 
  
```

number of visits at different times of the day    
* showing example for UFPs since all pollutant trend were similar    
* sites were visited every season and day of the week    
* sites were more likely to be visited during certain hours: early morning, early afternoon, and evening hours


```{r}

print("total number of site visits")

stops %>%
  #show UFPs as an example
  filter(variable == "pnc_noscreen") %>%

  gather(temporal_var, temporal_value, season, day, hour) %>%
  mutate(
    temporal_var = factor(temporal_var, levels = c("season", "day", "hour")),
    temporal_value = factor(temporal_value, 
                            levels = c(levels(stops$season), levels(stops$day), c(0:23) )
                              )
    ) %>%  
  # group_by(temporal_var, temporal_value, location, location_no, variable) %>%
  # summarize(count = n()) %>%
  
  ungroup() %>%
  mutate(location_no = factor(location_no)) %>%
  
  ggplot(aes(x=temporal_value,
             y=location_no,
             #group=location_no
             )) + 
  facet_grid(pollutant~temporal_var, scales="free") + 
  geom_bin2d(
    #binwidth = c(1, 1),
    #bins=length(unique(stops$location_no))
             ) +
  #scale_y_discrete(breaks = seq(10, 309, 10)) +
  scale_y_discrete(guide = guide_axis(check.overlap = T))+
  
  #scale_x_discrete(guide = guide_axis(check.overlap = T))+
  labs(
       x = "Sampling Time" ,
       y="Site",
       fill = "No. Visits"
       )

ggsave(file = file.path(image_path, "SI", "samples_per_site.png"), height = 9, width = 10)

```



## Concentration Table

```{r}

location_quant <- stops %>%
  #calc stats for alternative boxplots
  group_by(pollutant, ufp_range_nm, location, location_no) %>%
  alt_boxplot(var = "value", min_q = low_boxplot_q, max_q = high_boxplot_q) %>%
  mutate(iqr = Q75-Q25,
         qmax_qmin = Qmax/Qmin,
         max_min = Max/Min
         )

 

```



```{r}
# conc table

stops %>%
  left_join(instruments) %>%
  group_by(Pollutant=pollutant, Instrument) %>%
  summarize(
    N = n(),
    #Min = min(value),
    Q05 = quantile(value, 0.05),
    Q25 = quantile(value, 0.25),
    Median = quantile(value, 0.50),
    Mean = mean(value),
    Q75 = quantile(value, 0.75),
    Q95 = quantile(value, 0.95)
    #Max = max(value),
    ) %>%
  # mutate(
  # 'Q95/Q05' = format(Q95/Q05, digits = 2),
  # ) %>%
  mutate_at(
    vars(N:Q95), ~ifelse(grepl("PNC|CO2|BC", Pollutant), 
                           format(., digits=0, scientific = F, big.mark = ","),
                           format(., digits=2, scientific = F, big.mark = ",")
                           )
  ) %>%
  mutate(
    Instrument = factor(Instrument, levels = c("P-TRAK", "Screened P-TRAK", "P-TRAK, Screen", "NanoScan", "DiSCmini", "MA200","CAPS", "M903", "Li-Cor",  "-", "NA"))

  ) %>%
  arrange(Instrument) %>%
  kable(caption = "Distribution of site visit pollutant concentrations (N = 309 sites x ~ 29 visits/site).") %>%
  kable_styling()

```


## Site Variability

* boxplots of site visit medians.    
* whiskers are for the quantiles: `r low_boxplot_q` and `r high_boxplot_q`

```{r}
# load between site variabiltiy (later)

annual_conc1 <- readRDS(file.path("Data", "Output", "annual_conc_table.rda")) %>%
  mutate(
    IQR = Q75-Q25
  ) %>%
  select(Pollutant, Instrument, IQR) %>%
  mutate(
    IQR = ifelse(grepl("PNC|CO|BC|NO2", Pollutant), 
                           format(IQR, digits=0, scientific = F, big.mark = ",", 
                                  ),
                           format(IQR, digits=1, scientific = F, big.mark = ",", 
                                  )
                  )
    )


```


```{r, eval=F}

# # IQR table
# location_quant %>%
#   mutate(
#     pollutant = as.character(pollutant),
#     pollutant = ifelse(grepl("UFP", pollutant), paste0(pollutant, ", ", ufp_range_nm), pollutant)
#          ) %>%
#   #left_join(instruments, by=c("pollutant" = "pollutant2")) %>%
#   group_by(Pollutant=pollutant, #Instrument
#            ) %>%
#   summarize(
#     Min = min(iqr),
#     Median = quantile(iqr, 0.50),
#     Max = max(iqr),
#     ) %>%
#   mutate(
#     'Max/Min' = format(Max/Min, digits = 2),
#   ) %>%
#   mutate_at(
#     vars(Min:Max), ~ifelse(grepl("UFP|CO2|BC", Pollutant), 
#                            format(., digits=0, scientific = F, big.mark = ",",  
#                                   justify = "left"
#                                   ),
#                            format(., digits=2, scientific = F, big.mark = ",", 
#                                   justify = "left"
#                                   )
#                            )
#   ) %>% 
#   
#   mutate(
#     Min = gsub(" ", "", Min),    
#     Max = gsub(" ", "", Max),
#     Median = paste0(Median, " (", Min, "-", Max, ")")
#         ) %>%
#   select(
#     -c(Min, Max)
#   ) %>%
#   
#   left_join(annual_conc1) %>%
#   relocate(Instrument, .after=Pollutant) %>%
#   
#   kable(caption = 
#           "Comparison of within and between site variability. Within site variability is for site-specific IQR values of median visit concentrations (N = 309 site IQRs, each site has approximately 29 visits) and thus shows visit concentration data. Between site variability  is for the IQR of annual average concentrations (N=309 sites), as shown in Table ___.",
#         
#           #"Distribution of site-specific IQR values for median visit concentrations (N = 309 site IQRs, each site has approximately 29 visits).", 
#         col.names = c("",  "",
#                       "Median (Range) IQR", 
#                       "Max/Min IQR", 
#                       "IQR"), 
#         align = "r",   
#         #format.args = list(big.mark = ",")
#         ) %>%
#   kable_styling() %>%
#   add_header_above(c("Pollutant", "Instrument", "Within-Site Variability of Median Visit Conc's" = 2, "Between-Site Variability of Annual Avg Conc's" = 1))

```

 


boxplots

```{r}


# like table above, but formatted differently
visit_conc <- stops %>%
  group_by(pollutant, ufp_range_nm) %>%
  summarize(
    Q05 = quantile(value, 0.05),
    Median = quantile(value, 0.50),
    Q95 = quantile(value, 0.95),
    )  
 
```

 
```{r}
line_col <- "red"
line_a <- 0.5

```

 
 
 
 

 

## Temporal Variability 

notes   
* note, plotting non-extreme values, otherwise can't see any trends in the plots    
  - if on log scale, values <=0 are dropped (e.g., many for BC since aethalometer instrument is noisy)   
* note that different sites were visited at different times, so there could be some spatial confounding   
  

* NanoScan (10-420 nm) has higher readings than discmini (10-700 nm) b/c it is more efficient at measuring the smallest particles due to the difference in technologies. The Nanoscan measures optical diameter after alcohol condensation to grow the particles, whereas the disk-mini measures electrical mobility diameter based on the current produced when the particles are captured on a oppositely charged surface.  The electrostatic method has lower detection efficiencies for these very small particles than does the condensation/optical method.

by season 


```{r, fig.height=12}
# by month

print("Median site visit concentrations by season. Lines connect median monthly concentrations.")

stops %>%
    mutate(month =month(time, label = T),
            ufp_instrument = ifelse(ufp_instrument == "-", "NA", as.character(ufp_instrument)),
           ufp_instrument = factor(ufp_instrument, levels = c( "P-TRAK", "Screened P-TRAK", "NanoScan", "DiSCmini", "-", "NA"))
           ) %>%  
    #calc stats for alternative boxplots
    group_by(pollutant, ufp_range_nm, ufp_instrument, season) %>%
    alt_boxplot(var = "value", min_q = low_boxplot_q, max_q = high_boxplot_q) %>% #View()
    
    ggplot(aes(x=season, fill=ufp_instrument)) +
    geom_boxplot(aes(ymin=Qmin, lower=Q25, middle=Q50, upper=Q75, ymax=Qmax), stat = "identity") + 
  facet_wrap(~pollutant, 
             scales="free", nrow = 3) +
    labs(y= "Value",
         x="Season",
         fill = "PNC Instrument"
         )  

ggsave(file = file.path(image_path, "SI", "temporal_season.png"), height = 8, width = 7)

```


by weekday

```{r, fig.height=12}
print("Median site visit concentrations by day of the week. Lines connect median monthly concentrations.")

stops %>%
  mutate(dow = ifelse(day %in% c("Sat", "Sun"), "Weekend", "Weekday"),
         ufp_instrument = ifelse(ufp_instrument == "-", "NA", as.character(ufp_instrument)),
           ufp_instrument = factor(ufp_instrument, levels = c( "P-TRAK", "Screened P-TRAK", "NanoScan", "DiSCmini", "-", "NA"))
         ) %>%
  #calc stats for alternative boxplots
  group_by(pollutant, ufp_range_nm, ufp_instrument, dow) %>%
  alt_boxplot(var = "value") %>%  
  
  ggplot(aes(x=dow, fill=ufp_instrument)) +
  geom_boxplot(aes(ymin=Qmin, lower=Q25, middle=Q50, upper=Q75, ymax=Qmax), stat = "identity") + 
  facet_wrap(~pollutant, scales="free", nrow = 3) +
  labs(y= "Value", x="Day",
       fill = "PNC Instrument"
       )

ggsave(file = file.path(image_path, "SI", "temporal_day.png"), height = 8, width = 7)                 
```

by hour 

```{r, fig.height=12}

stops %>%
  #don't include midnight, since little data here, and it makes it seem like instrument clocks were off
  filter(hour !=0) %>%
  mutate(hour = factor(hour, levels = c(4:23))) %>%
    #calc stats for alternative boxplots
    group_by(pollutant, ufp_instrument, hour) %>%
    alt_boxplot(var = "value") %>%   
    
    ggplot(aes(x=hour, fill=ufp_instrument)) +
    geom_boxplot(aes(ymin=Qmin, lower=Q25, middle=Q50, upper=Q75, ymax=Qmax), stat = "identity") + 
     
  facet_wrap(~pollutant+ufp_instrument, scales="free", nrow = 4) +
  theme(legend.position = "none") +
    
    labs(y= "Value",
         x= "Hour" 
         )  

ggsave(file = file.path(image_path, "SI", "temporal_hour.png"), height = 8, width = 8)                 
```

spaghetti plots
 
```{r}
"Site-specific pollutant concentrations over. Thin lines show site-specific smooth (loess) fits for median visit concentrations (N~29 visits/site). Black lines show the overall smooth trends for all of the sites."

stops %>%
  mutate(#ufp_range_nm = ifelse(ufp_range_nm == "-", "NA", ufp_range_nm)
         ) %>%
    ggplot(aes(x=time, y=value)) +
       geom_smooth( se=F, size=0.1, alpha=0.8,
                    aes(col=ufp_instrument, group=location), show.legend = F) +
  #overall trend
  geom_smooth(se=F, col="black") +
  
  facet_wrap(~pollutant+ufp_instrument, scales="free", nrow = 4) +
    labs(y= "Value",
         x= "Date",
         col = "PNC Instrument"
         )  

ggsave(file = file.path(image_path, "SI", "spaghetti_plots.png"), height = 9, width = 9)                 
```


 
## ANOVA


```{r}
# using all of the observations to characterize variability 
var_df <- data.frame()

for(i in unique_pollutants2) {
  # i=unique_pollutants2[1]
  temp <- anovaVCA(value ~ location + day+ hour + season,
           Data = as.data.frame(subset(stops, pollutant2 == i) ))$aov.tab  %>% # 
    tidy() %>%
    select(Variable = .rownames, DF, Percent = X.Total) %>%
    mutate(DF = round(DF))
  
  names(temp)[names(temp) == "Percent"] <- i
  
  if(i==unique_pollutants2[1]) {
    var_df <- temp
    } else {
                                    #only use DF from the first pollutant, as an exp
      var_df <- left_join(var_df, select(temp, -DF))
    }
  }

 
```

```{r}
var_df %>%
  gather(pollutant, percent, 'BC (ng/m3)':'PNC (pt/cm3), Screened P-TRAK') %>%
  filter(Variable != "total") %>%
  mutate(
    variability = ifelse(Variable == "location", "spatial",
                  ifelse(Variable == "error", "residual", "temporal")
                  ),
    variability = factor(variability, levels = c("residual", "temporal", "spatial"))
  ) %>%
  group_by(variability, pollutant) %>%
  summarize(
    percent = sum(percent)
  ) %>%
  spread(pollutant, percent) %>%
  kable(caption = "total variability categorized", digits = 0) %>%
  kable_styling()



```


plot


```{r}
var_df %>%
  gather(pollutant, percent, 'BC (ng/m3)':'PNC (pt/cm3), Screened P-TRAK') %>%
  filter(Variable != "total") %>%
  mutate(
    Variable = factor(Variable, levels = c("error", "location", "season", "day", "hour")),
    variability = ifelse(Variable == "location", "spatial",
                  ifelse(Variable == "error", "residual", "temporal")
                  ),
    variability = factor(variability, levels = c("residual", "temporal", "spatial")),
    pollutant = factor(pollutant, levels =
                         c("PNC (pt/cm3), P-TRAK", "PNC (pt/cm3), Screened P-TRAK", "PNC (pt/cm3), NanoScan", "PNC (pt/cm3), DiSCmini", "NO2 (ppb)", "BC (ng/m3)", "PM2.5 (ug/m3)", "CO2 (ppm)"
                           )
                       )
    ) %>%
  ggplot(aes(x=variability, y=percent, fill=Variable)) + 
  facet_wrap(~pollutant) + 
  geom_bar(stat = "identity") + 
  labs(fill = "",
       x = "Variability", y = "Percent (%)"
       ) 


ggsave(file = file.path(image_path, "SI", "anova.png"), height = 9, width = 9)                 

```




 
 

 

# Annual averages

* calculating a simple mean annual average using all the available site data 


## Concentration Table

```{r}
annual <- stops %>%
  left_join(instruments) %>% #View()
  group_by(variable, pollutant, ufp_instrument, Instrument, location) %>%
  summarize(value = mean(value)) %>%
  ungroup() %>%
  mutate(
    pollutant2 = ifelse(grepl("PNC", pollutant), paste0(pollutant, ", ", ufp_instrument), as.character(pollutant ))
  ) %>%
  #add location lat/long
  left_join(locations)

```


```{r}
#table of distribution
annual_conc <- annual %>%
  mutate(
    pollutant = as.character(pollutant),
    pollutant = ifelse(grepl("PNC", pollutant), paste0(pollutant, ", ", ufp_instrument), pollutant)
         ) %>%
  #left_join(instruments, by = c("pollutant" = "pollutant2")) %>%
  group_by(Pollutant=pollutant, Instrument
           ) %>%
  summarize(
    Min = min(value),
    Q05 = quantile(value, 0.05),
    Q25 = quantile(value, 0.25),
    Median = quantile(value, 0.50),
    Q75 = quantile(value, 0.75),
    Q95 = quantile(value, 0.95),
    Max = max(value),
    ) %>%
  mutate(#'Q95/Q05' = round(Q95/Q05, 1),
         'Max/Min' = round(Max/Min, 1),
         #'IQR/Median' = round((Q75-Q25)/Median, 1)
         ) 

# save for use above. add this between site variability to within site variability table above
saveRDS(annual_conc, file = file.path("Data", "Output", "annual_conc_table.rda"))

annual_conc %>%
  mutate_at(
    vars(Min:Max), ~ifelse(grepl("PNC|CO|BC", Pollutant), 
                           format(., digits=0, scientific = F, big.mark = ","),
                           format(., digits=2, scientific = F, big.mark = ",")
                           #round(.,), 
                           #round(.,1)
                           )
  ) %>%
  
  kable(caption = "Distribution of annual average pollutant concentrations at monitoring sites (N = 309)") %>%
  kable_styling()
  
```

 
```{r}
print("Annual average site concentrations")
  
#boxplots
annual %>%
    mutate(#ufp_instrument = ifelse(ufp_instrument == "-", "NA", as.character(ufp_instrument)),
           
            ufp_instrument = ifelse(ufp_instrument == "-", "NA", as.character(ufp_instrument)),
           ufp_instrument = factor(ufp_instrument, levels = c( "P-TRAK", "Screened P-TRAK", "NanoScan", "DiSCmini", "-", "NA")),
           ) %>%

  ggplot(aes(x=pollutant, y=value, fill=ufp_instrument)) + 
  facet_wrap(~pollutant, scales= "free") + 
  geom_boxplot() +
  labs(x="Pollutant",
       y = "Value",
       fill = "PNC Instrument"
       )

ggsave(file.path(image_path, "SI", "annual_avg.png"), height = 6, width = 7)

```



## Pollutant comparisons

```{r}

cor_df <- annual %>%
  mutate(pollutant = if_else(grepl("PNC", pollutant), 
                            paste0(pollutant, "\n", ufp_instrument),
                            as.character(pollutant)
                            )
         ) %>%
  select(pollutant, location, value) %>%  
  spread(pollutant, value) %>% 
  select(-location) 
  
```

```{r}
#function colors upper panel by correlation value
# source: https://stackoverflow.com/questions/45873483/ggpairs-plot-with-heatmap-of-correlation-values 

color_cor <- function(data, mapping, method="p", use="pairwise", alpha=0.5, ...){

              # grab data
              x <- eval_data_col(data, mapping$x)
              y <- eval_data_col(data, mapping$y)

              # calculate correlation
              corr <- cor(x, y, method=method, use=use)

              # calculate colour based on correlation value
              # Here I have set a correlation of minus one to blue, 
              # zero to white, and one to red 
              # Change this to suit: possibly extend to add as an argument of `my_fn`
              colFn <- colorRampPalette(c("blue", "white", "red"), interpolate ='spline')
              fill <- colFn(100)[findInterval(corr, seq(-1, 1, length=100))]

              ggally_cor(data = data, mapping = mapping, 
                         #don't include stars if correlations "sign"
                         stars = FALSE,
                         digits=2,
                         ...) + 
                theme_void() +
                theme(panel.background = element_rect(fill=alpha(fill, alpha)))
            }
```


```{r, fig.height=12, fig.width=10}
print("Annual average pollutant correlations")

# correlation plot
ggpairs(cor_df, 
        upper = list(continuous = color_cor),
        lower = list(continuous = wrap("smooth_loess", alpha=0.2)),
        ) + #scale_x_continuous(labels = scales::scientific) 
  scale_x_continuous(guide = guide_axis(check.overlap = TRUE))

ggsave(file.path(image_path, "Other", "correlations_estimtates.png"), height = 11, width = 11)

```

 


## Maps 


```{r}
#  
annual2 <- st_as_sf(annual, coords = c("longitude", "latitude"),  remove = F, crs=crs_deg) #%>% st_transform(crs = crs_m)

unique_p <- distinct(annual[c("pollutant", "ufp_instrument")])

```


```{r, fig.height=18}
#non-UFPs. each pollutant has its own legend

print("Annual average pollutant concentrations at monitoring sites (N=309).")

p <- list() 

for(i in seq_along(non_ufp_p)) {
   #i=1
   
  p[[i]] <- ggmap(ggmap = map0, darken = c(.5, "white")) +
    #monitoring area
    geom_sf(data=monitoring_area_shp, inherit.aes = F, aes(fill = "Monitoring Area"),
            alpha=0.1,
            #reduce/eliminate outline
            lwd = 0.1,
            ) +
    
    #annual averages
    geom_point(data=filter(annual, pollutant == non_ufp_p[i]),
             aes(x=longitude, y=latitude, col=value), 
             inherit.aes = F, size=2)  +
    scale_color_gradient(name = "Conc",
                        low = "yellow", high = "red",
                        ) +
    facet_wrap(~pollutant)  + 
    # add scale & N arrow to top right
    annotation_scale(location = "tr") +
    annotation_scale(location = "tr", unit_category ="imperial", pad_y = unit(0.55, "cm")) +

    annotation_north_arrow(location = "tr",
                           #point towards North Pole
                           which_north = "true",
                           pad_y = unit(0.5, "in"),
                           style = north_arrow_fancy_orienteering
                           ) +
  theme_bw() +
  theme(
    legend.justification=c(0,1),  
    legend.position=c(0,1),  
    legend.background =  element_blank()
    ) +
    coord_sf(expand = F) +
    scale_x_continuous(breaks = map_x_labels, labels = map_x_labels ) +
    scale_y_continuous(breaks = map_y_labels, 
                       #labels = map_y_labels
                       labels = format(map_y_labels,digits = 1,
                                       nsmall = 1
                                       )
                       ) +
    #arrange legend order
  guides(col = guide_colorbar(order = 2),
              fill = guide_legend(order = 1)
         ) +
    #add attribution/reference to bottom left
    geom_text(aes(x=-Inf, y=-Inf, hjust=-0.01, vjust=-0.3,
                label= "Map tiles by Stamen Design, under CC BY 3.0. \nData by OpenStreetMap, under ODbL."
                ),
            size=2.5
            ) +
  labs(x = "Longitude",
       y = "Latitude",
       fill = ""
       ) 
  
  p[[i]]
  
  }

ggarrange(plotlist = p, ncol = 2, nrow = 2)

ggsave(file.path(image_path, "SI", "map_non_ufp.png"),  height = 13, width = 11)

```

* UFP predictions seem to vary depending on what instrument was used?    
  * note, people wouldn't normally use screened PTRAKS (36-1,000 nm particles) to measure "UFPs"


-all UFP instruments (for SI)
  

```{r}
# ufp_instruments

lapply(ufp_instruments, function(x){
  # x = ufp_instruments[1]
  
  ggmap(ggmap = map0, darken = c(.5, "white")) +
    #monitoring area
    geom_sf(data=monitoring_area_shp, inherit.aes = F, aes(fill = "Monitoring Area"),
            alpha=0.1, lwd = 0.1) +
    
    #annual averages
    geom_point(data= filter(annual, pollutant == ufp_p, ufp_instrument == x),
             aes(x=longitude, y=latitude, col=value), inherit.aes = F, size=2)  +
  scale_color_gradient(name = "Conc", low = "yellow", high = "red") +
    facet_wrap(~pollutant+ufp_instrument)  + 
    # add scale & N arrow to top rught
    annotation_scale(location = "tr") +
    annotation_scale(location = "tr", unit_category ="imperial", pad_y = unit(0.55, "cm")) +

    annotation_north_arrow(location = "tr",
                           #point towards North Pole
                           which_north = "true",
                           pad_y = unit(0.5, "in"),
                           style = north_arrow_fancy_orienteering
                           ) +
  theme_bw() +
  theme(
    legend.justification=c(0,1),  
    legend.position=c(0,1),  
    legend.background =  element_blank(),
    ) +
    coord_sf(expand = F) +
    scale_x_continuous(breaks = map_x_labels, labels = map_x_labels ) +
    scale_y_continuous(breaks = map_y_labels, 
                       labels = format(map_y_labels,digits = 1, nsmall = 1)
                       ) +
  #arrange legend order
  guides(col = guide_colorbar(order = 2),
              fill = guide_legend(order = 1)) +
  #add attribution/reference to bottom left
    geom_text(aes(x=-Inf, y=-Inf, hjust=-0.01, vjust=-0.3,
                label= "Map tiles by Stamen Design, under CC BY 3.0. \nData by OpenStreetMap, under ODbL."
                ),
            size=2.5
            ) +
  labs(x = "Longitude",
       y = "Latitude",
       fill = ""
       )
  }) %>%
  ggarrange(plotlist = .)

ggsave(file.path(image_path, "SI", "map_ufp.png"),  height = 13, width = 11)

```


```{r, fig.height=18, eval=F}
#UFPs. these share the same legend

ggmap(ggmap = map0, darken = c(.5, "white")) +
    #monitoring area
    geom_sf(data=monitoring_area_shp, inherit.aes = F, aes(fill = "Monitoring Area"),
            alpha=0.1, lwd = 0.1) +
    
    #annual averages
    geom_point(data=filter(annual, pollutant == ufp_p),
             aes(x=longitude, y=latitude, col=value), 
             inherit.aes = F, size=2,
             )  +
  scale_color_gradient(name = "Conc",
                       low = "yellow", high = "red",
                       ) +
    facet_wrap(pollutant+ufp_instrument, nrow = 2)  + 
    # add scale & N arrow to top rught
    annotation_scale(location = "tr") +
    annotation_scale(location = "tr", unit_category ="imperial", pad_y = unit(0.55, "cm")) +

    annotation_north_arrow(location = "tr",
                           #point towards North Pole
                           which_north = "true",
                           pad_y = unit(0.5, "in"),
                           style = north_arrow_fancy_orienteering
                           ) +
  theme_bw() +
  theme(
    legend.justification=c(0,1),  
    legend.position=c(0,1),  
    legend.background =  element_blank(),
    ) +
    coord_sf(expand = F) +
    scale_x_continuous(breaks = map_x_labels, labels = map_x_labels ) +
    scale_y_continuous(breaks = map_y_labels, 
                       labels = format(map_y_labels,digits = 1, nsmall = 1)
                       ) +
  #arrange legend order
  guides(col = guide_colorbar(order = 2),
              fill = guide_legend(order = 1)) +
  #add attribution/reference to bottom left
    geom_text(aes(x=-Inf, y=-Inf, hjust=-0.01, vjust=-0.3,
                label= "Map tiles by Stamen Design, under CC BY 3.0. \nData by OpenStreetMap, under ODbL."
                ),
            size=2.5
            ) +
  labs(x = "Longitude",
       y = "Latitude",
       fill = ""
       ) 

ggsave(file.path(image_path, "SI", "map_ufp.png"),  height = 13, width = 11)

```


  


  




```{r} 
# Session Info
# # sessionInfo()

```
 
 
```{r, eval=F}
# save package references to reference manager #https://bookdown.org/yihui/rmarkdown-cookbook/write-bib.html
##1.
knitr::write_bib(c("broom", "colorspace", "cowplot", "fmsb", "forcats", "GGally", "ggpmisc", "ggpp", "ggpubr", "gstat", "knitr", "spData", "units" 
                   ), 
                 file = file.path("Data", "Output", "packages_v2.bib"))
##2. import this file into reference manager 

```




