---
title: 'Aim 2: Discussion'
author: "Magali Blanco"
date: '`r Sys.Date()`'
output:
  html_document:
    number_sections: yes
    df_print: paged
    toc: yes
    toc_depth: '3'
editor_options:
  chunk_output_type: console
---


**TO DO**   


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = F, 
                      cache=F, cache.comments = F, 
                      message = F, warning = F, 
                      tidy.opts=list(width.cutoff=60), tidy=TRUE,
                      dpi=300,
                      fig.width=8, 
                      fig.height=6
                      )  
set.seed(1)


pacman::p_load(knitr, kableExtra,
               tidyverse, dplyr,
               ggrepel #avoid overlapping labels when plotting
               )  

theme_set(theme_bw())
theme_update(legend.position = "bottom")

```

# Load data

```{r}
# model performance
## ufp
ufp_r2 <- readxl::read_xlsx(file.path("..", "Manuscript", "Mean UFP & BC comparison.xlsx"), sheet = 3) %>%
  #only include studies that have been fully entered into excel or qualify
  filter(Include_in_analysis == TRUE) %>% select(-Include_in_analysis) %>%
  mutate(
    Mean_Model_R2 = as.numeric(Mean_Model_R2),
    #shorten names
    Location = gsub("Amsterdam", "AMS", Location),
    Location = gsub("Rotterdam", "RTM", Location),
    #Sampling_type = factor(Sampling_type, levels = c("Long-term Stationary", "Short-term Stationary", "Short-term Non-stationary"))
  ) %>%
  arrange(Sampling_type, Study) %>%
  #remove white spaces from start & end of strings
  mutate_if(is.character, ~str_trim(.))

## bc
bc_r2 <- readxl::read_xlsx(file.path("..", "Manuscript", "Mean UFP & BC comparison.xlsx"), sheet = 4) %>%
   #only include studies that have been fully entered into excel or qualify
  filter(Include_in_analysis == TRUE) %>% select(-Include_in_analysis) %>%
  mutate(
    #only keep few digits - for some reason some values have many decimal #s
    Model_R2 = ifelse(nchar(Model_R2) > 9, substr(Model_R2, 1,4), Model_R2),
    Mean_Model_R2 = as.numeric(Mean_Model_R2),
    #shorten names
    Location = gsub("Amsterdam", "AMS", Location),
    Location = gsub("Rotterdam", "RTM", Location),
    
   # Sampling_type = factor(Sampling_type, levels = c("Long-term Stationary", "Short-term Stationary", "Short-term Non-stationary"))
    ) %>%
  arrange(Sampling_type, Study)%>%
  #remove white spaces from start & end of strings
  mutate_if(is.character, ~str_trim(.))

common_r2_names <-str_subset(string = names(bc_r2),pattern =  paste0(names(ufp_r2), collapse = "|"), )   

```


```{r}
#concentrations
## ufp
ufp0 <- readxl::read_xlsx(file.path("..", "Manuscript", "Mean UFP & BC comparison.xlsx")) %>%
  #only include studies that have been fully entered into excel or qualify
  filter(Include_in_analysis == TRUE) %>%
  select(-Include_in_analysis) %>%
  #use median if mean is missing
  mutate(Mean = ifelse(is.na(Mean), yes = Median, no = Mean),
         Location = gsub("Amsterdam", "AMS", Location),
         Location = gsub("Utrecht", "UT", Location),
         Location = gsub("Maastricht", "MMa", Location),
         Location = gsub("Rotterdam", "RTM", Location),

         This_study = grepl("Blanco", Study),
         
         Year = as.numeric(str_extract(Study, "[0-9]{4}")),
         
         #Sampling_type = factor(Sampling_type, levels = c("Long-term Stationary", "Short-term Stationary", "Short-term Non-stationary"))
         ) %>%
  #remove white spaces from start & end of strings
  mutate_if(is.character, ~str_trim(.))

ours <- ufp0 %>%
  filter(grepl("Blanco", Study))
 
## bc
bc0 <- readxl::read_xlsx(file.path("..", "Manuscript", "Mean UFP & BC comparison.xlsx"), sheet=2) %>%
  #only include studies that have been fully entered into excel or qualify
  filter(Include_in_analysis == TRUE) %>%
  select(-Include_in_analysis) %>%
  mutate(
         Location = gsub("Amsterdam", "AMS", Location),
         Location = gsub("Utrecht", "UT", Location),
         Location = gsub("Maastricht", "MMa", Location),
         Location = gsub("Rotterdam", "RTM", Location),
         
         This_study = grepl("Blanco", Study),
         
         #Sampling_type = factor(Sampling_type, levels = c("Long-term Stationary", "Short-term Stationary", "Short-term Non-stationary"))
         ) %>%
  rename(Mean = Mean_Median) %>%
  #remove white spaces from start & end of strings
  mutate_if(is.character, ~str_trim(.))

bc_ours <- bc0 %>%
  filter(grepl("Blanco", Study))

common_conc_names <-str_subset(string = names(bc0),pattern =  paste0(names(ufp0), collapse = "|"), )   
```

 
```{r}
#common variables
text_size <- 2.5
image_path <- file.path("..", "Manuscript", "Images")
```

# Model R2

## Plots


-Note: plotting the primary analysis R2 or the average R2 if multiple models were run without a clear indication of which was the primary model

- Our study model performance is the avg R2 for the primary analysis, using 10-fold CV and test site predictions.


```{r}
"Model R2 estimates from other studies"

ufp_r2 %>%
  select(common_r2_names) %>%
  rbind(bc_r2[common_r2_names]) %>% 
  mutate(Pollutant = c(rep("PNC", nrow(ufp_r2)), rep("BC", nrow(bc_r2)) ),
         Pollutant = factor(Pollutant, levels = c("PNC", "BC")),
    this_study = ifelse(grepl("Blanco", Study), "This Study", "Other Study" ),
    Sampling_type = gsub(" ", "\n", Sampling_type)
    ) %>% 
  group_by(Pollutant) %>%
  mutate(this_study_r2 = Mean_Model_R2[grepl("Blanco", Study)]) %>%   
  
  ggplot(data=., aes(x=Sampling_type, y=Mean_Model_R2, label = Location,)) +
  geom_line(alpha=0.4) +
  #stat_summary(fun.y = "median", geom = "point", size = 3, alpha=0.4) +
  geom_point(aes(col = this_study,), size=2) +
  
  # #city names 
  geom_text_repel(aes(col = this_study,), size = text_size)    +
  #this study
  geom_hline(aes(yintercept = this_study_r2), linetype = "dashed", alpha=0.5) +
  facet_wrap(~Pollutant ) +
  theme(plot.caption = element_text(size=8, hjust = 0), legend.position = "bottom") +
  labs(y = "Model R2",
       x = "Sampling Type",
       caption = paste0("AMS = Amsterdam; UT = Utrecht"),
       col = ""
       ) 
  
ggsave(file.path(image_path, "discussion_r2.png"), width = 7, height=6)

```

-just UFP as example & larger text

```{r, eval=F}
"Model R2 estimates from other studies"

ufp_r2 %>%
  select(common_r2_names) %>%
  rbind(bc_r2[common_r2_names]) %>% 
  mutate(Pollutant = c(rep("PNC", nrow(ufp_r2)), rep("BC", nrow(bc_r2)) ),
         Pollutant = factor(Pollutant, levels = c("PNC", "BC")),

    # are they calculating a site average from >1 visit & using this in their analysis (e.g., vs using individual site visits in a regression)
    multi_visit_mean = visits_per_site > 1 & site_mean_only == "yes",
    multi_visit_mean = ifelse(is.na(multi_visit_mean), "unclear", multi_visit_mean),

    #this_study = grepl("Blanco", Study),
    this_study = ifelse(grepl("Blanco", Study), "This Study", "Other Study" ),

    
  ) %>% 
  group_by(Pollutant) %>%
  mutate(this_study_r2 = Mean_Model_R2[grepl("Blanco", Study)]
         
         ) %>% 
  
  #just UFP for example ppt
  filter(Pollutant == "PNC") %>%
  
  
  ggplot(data=., aes(x=Sampling_type, y=Mean_Model_R2,
             label = Location,
             )) + 
  
      geom_line(alpha=0.4) +
     stat_summary(fun.y = "median", geom = "point", size = 4, 
                  alpha=0.4
                  ) +
      
      geom_point(aes(#shape=multi_visit_mean,
             col = this_study,),
             size=3
             ) +
  
  # #city names 
  geom_text_repel(aes(col = this_study,), size = text_size+1,  
                  #fontface = "plain",  
                    )    +
  #this study
  geom_hline(#yintercept = .$this_study_r2, #.$Mean_Model_R2[.$this_study],
    aes(yintercept = this_study_r2),
             linetype = "dashed",
             alpha=0.5
             ) +

      facet_wrap(~Pollutant ) +
      
      theme(plot.caption = element_text(size=8,
                                    hjust = 0),
            legend.position = "bottom",
            
            #make general text larger
            text = element_text(size=20)
            ) +
  #avoid overlapping x axis ticks
  #scale_x_discrete(labels = function(x) str_wrap(x, width = 10))+
  ##scale_x_discrete(guide = guide_axis(n.dodge=2)) +

  labs(#title = "Model R2 estimates from other studies",
       y = "Model R2",
       x = "Sampling Type",
       caption = paste0("AMS = Amsterdam; UT = Utrecht \nGray dot is the median; vertical line is the range"),
       col = "" #"This Study",
       #shape = "Multi-site \nAvg"
       ) 

```


### sampling approaches in other studies (all)

**UFP**

```{r}
"Sampling approaches across other UFP studies"

ufp_r2 %>%
  mutate(Sampling_type = gsub(" ", "\n", Sampling_type)) %>%  
  gather("variable", "Value", visits_per_site, visit_duration_min, total_site_duration_min, sites_per_model) %>%
  mutate(This_study = ifelse(grepl("Blanco", Study), "This Study", "Other Study"),
         
         variable = recode_factor(factor(variable),
                                  "sites_per_model" = "sites",
                                  "visit_duration_min" = "visit duration (min)",
                                  "total_site_duration_min" = "total site duration (min)",
                 "visits_per_site"  = "visits per site"               
         ),
         variable = factor(variable, levels = c("sites", "visits per site", "visit duration (min)","total site duration (min)")
                           ),
         ) %>%
  {
  ggplot(data=., aes(x=Sampling_type, y=Value)) +
      #stat_summary(fun.y = "median", geom = "point", size = 3,  alpha=0.4) +
      geom_point(aes(col = This_study)) +
      #squares around our study
      geom_point(data = filter(., This_study=="This Study"), aes(), shape=22, size=5, alpha=0.8) +
  
      #avoid overlapping x axis ticks
      #scale_x_discrete( labels = function(x) str_wrap(x, width = 15))+
      scale_y_log10() +
      facet_wrap(~variable, scales = "free") +
      theme(legend.position = "bottom") +
      labs(x = "Sampling type", 
           y= "",
           col = ""  
           )
  }


ggsave(file.path(image_path, "SI", "sampling_approach_ufp.png"), width = 7, height=6)

```



```{r, eval=F}
# -example w/ just top 2 rows, larger text

"Sampling approaches across other UFP studies"

ufp_r2 %>%
 
  mutate(
    multi_visit_mean = visits_per_site > 1 & site_mean_only == "yes",
    multi_visit_mean = ifelse(is.na(multi_visit_mean), "unclear", multi_visit_mean),
  ) %>%  
  gather("variable", "Value", visits_per_site, visit_duration_min, 
         #total_site_duration_min, 
         sites_per_model
         ) %>%
  mutate(This_study = ifelse(grepl("Blanco", Study), "This Study", "Other Study"),
         
         variable = recode_factor(factor(variable),
                                  "sites_per_model" = "sites",
                                  "visit_duration_min" = "visit duration (min)",
                                  "total_site_duration_min" = "total site duration (min)",
                 "visits_per_site"  = "visits per site"               
         ),
         variable = factor(variable, levels = c("sites", "visits per site", "visit duration (min)","total site duration (min)")
                           ),
         ) %>% #View()
  
   #don't inlcude this since its incomplete
  filter( !(grepl("no stops", Sampling_type, ignore.case = T) & variable == "visit duration (min)")
          ) %>%
  {
  
  ggplot(data=., aes(x=Sampling_type, y=Value)) +
      
      #stat_summary(fun.y = "median", geom = "point", size = 4,  alpha=0.4) +

  geom_point(aes(col = This_study, 
                 ),
             size=3, 
             alpha=0.8
             ) +
  
  #squares around our study
  geom_point(data = filter(., This_study=="This Study"),
             aes(#col = This_study, 
                 ),
             shape=22,
             size=8, 
             alpha=0.8
             ) +
     
      #avoid overlapping x axis ticks
      scale_x_discrete(labels = function(x) str_wrap(x, width = 10))+
      scale_y_log10() +
      
      facet_wrap(~variable, scales = "free"
                 ) +
  
  theme(legend.position = "bottom",
        #make general text larger
        text = element_text(size=17),
        ) +
  
      labs(#title = "Sampling approaches across other UFP studies",
           x = "Sampling Type",
           col = "" #,
           #shape = "Multi-site \nAvg Modelled"
            
           )
}
```


**BC**

```{r}
"Sampling approaches across other BC studies"

bc_r2 %>%
  mutate(Sampling_type = gsub(" ", "\n", Sampling_type)) %>%  
  gather("variable", "Value", visits_per_site, visit_duration_min, total_site_duration_min, sites_per_model) %>%
  mutate(This_study = ifelse(grepl("Blanco", Study), "This Study", "Other Study"),
         
         variable = recode_factor(factor(variable),
                                  "sites_per_model" = "sites",
                                  "visit_duration_min" = "visit duration (min)",
                                  "total_site_duration_min" = "total site duration (min)",
                 "visits_per_site"  = "visits per site"               
         ),
         variable = factor(variable, levels = c("sites", "visits per site", "visit duration (min)","total site duration (min)")
                           )
         ) %>%
  {
  ggplot(data=., aes(x=Sampling_type, y=Value)) +
      #stat_summary(fun.y = "median", geom = "point", size = 3,alpha=0.4) +
      geom_point(aes(col = This_study)) +
     
  #squares around our study
  geom_point(data = filter(., This_study=="This Study"), shape=22, size=5, alpha=0.8) +
  
      #avoid overlapping x axis ticks
      # scale_x_discrete(#guide = guide_axis(n.dodge=2)
      #   labels = function(x) str_wrap(x, width = 15)
      #                  )+
      scale_y_log10() +
      facet_wrap(~variable, scales = "free") +
      theme(legend.position = "bottom") +
      labs(x = "Sampling type",
           col = ""
           )
    
  }

ggsave(file.path(image_path, "SI", "sampling_approach_bc.png"), width = 7, height=6)

```


## Tables

### --> update Sampling_type if change study order

**UFP**

```{r}
ufp_r2 %>%
  select(
    ## update if change study order!
    #Sampling_type,
    
    Study, Location, # = Study_location,
    Sampling = Sampling_approach, 
    
    Analysis_uses_site_mean = site_mean_only, Model_R2) %>%
  rename_if( grepl("_", names(.)), ~gsub("_", " ", .) ) %>%
  kable(caption = "Model R2 estimates from other LUR PNC studies", 
        digits = 2) %>%
  
  ### --> check that these are correct in the end
  pack_rows("Long-term Stationary", 1, 5) %>%
  pack_rows("Short-term Non-stationary", 6, 15) %>%
  pack_rows("Short-term Stationary", 16, 23) %>%
  add_footnote(., label = c("Location: AMS = Amsterdam; MMa = Maastricht; RTM = Rotterdam; UT = Utrecht"),
                notation = "none"
                ) %>%
  kable_styling()
  
```


**BC**

```{r}
bc_r2 %>%
 select(
   ## update if change study order!
   #Sampling_type,
   
   Study, Location,# = Study_location, 
   Sampling = Sampling_approach, Analysis_uses_site_mean = site_mean_only, Model_R2) %>%  
  rename_if( grepl("_", names(.)), ~gsub("_", " ", .) ) %>%
  kable(caption = "Model R2 estimates from other LUR BC studies", 
        digits = 2
        ) %>%
  pack_rows("Long-term Stationary", 1, 9) %>%
  pack_rows("Short-term Non-stationary", 10, 14) %>%
  pack_rows("Short-term Stationary", 15, nrow(bc_r2)) %>%
  add_footnote(., label = c("Location: AMS = Amsterdam; MMa = Maastricht; RTM = Rotterdam; UT = Utrecht"),
               notation = "none"
               ) %>%
  kable_styling()

```





```{r}
knitr::knit_exit()
```



# Concentrations in other studies

## Plots

both pollutants 

```{r, fig.height=7}
"Mean (or median) Conc reported by other studies"

set.seed(2)

ufp0 %>%
  select(common_conc_names) %>%
  rbind(bc0[common_conc_names]) %>%
  mutate(Pollutant = c(rep("PNC (pt/cm3)", nrow(ufp0) ), rep("BC (ng/m3)", nrow(bc0) )),
         Pollutant = factor(Pollutant, levels = c("PNC (pt/cm3)", "BC (ng/m3)"))
         ) %>%  
  
  ggplot(aes(y = Mean,
           x=Sampling_type, #x=1,
           )) + 
  
   geom_line(alpha=0.4) +
     stat_summary(fun.y = "median", geom = "point", size = 3, alpha=0.4) +
  
  geom_point(aes(col = This_study),
             #alpha=0.6
             ) + 
  # #city names 
  geom_text_repel(size = text_size, #-0.5, 
                  fontface = "plain",  
                  #make labels more spaced out
                  #box.padding = 0.5, 
                  aes(label = Location,
                      col = This_study
                      )
                   ) +
  
  # # Add a second axis and specify its features
  # scale_y_continuous(sec.axis = sec_axis(~./ours$Mean, 
  #                                        name="PNC Conc Relative to This Study", 
  #                                        breaks = seq(1,9, 2)
  #                                        )
  #                    ) +
  # #this study
  # geom_hline(yintercept = ours$Mean,
  #            linetype = "dashed",
  #            alpha=0.5
  #            ) +
  facet_wrap(~Pollutant, scales="free") +
    theme(legend.position = "bottom") +
  
  #avoid overlapping x axis ticks
  scale_x_discrete(guide = guide_axis(n.dodge=2)) +

  
  labs(x = "Sampling Type",
       y = "Mean Concentration",
       caption = "AMS = Amsterdam; MMa = Maastricht; RTM = Rotterdam; UT = Utrecht \nBlack line and dot are the range and median",
       col = "This Study"
       )


```


## Tables

**UFP**

```{r}
ufp0 %>%
  group_by(Study) %>%
   select(Study, Location,  
    Mean_UFP_Conc = Mean,
    Instrument, Instrument_range
  ) %>%
    
   rename_if( grepl("_", names(.)), ~gsub("_", " ", .) ) %>%
  kable(caption = "Average PNC Concentrations (pt/cm3) reported in other LUR PNC studies", 
        digits = 2) %>%
  kable_styling()
  
```

**BC** 

```{r}
bc0 %>%
  #group_by(Study) %>%
   select(Study, Location,  
    Mean_BC_Conc = Mean,
  ) %>%
   rename_if( grepl("_", names(.)), ~gsub("_", " ", .) ) %>%
  kable(caption = "Average BC Concentrations (ng/m3) reported in other LUR studies", 
        digits = 2) %>%
  kable_styling()
```


# UFP instrument ranges

- instrument ranges are not necessarily related to observed higher/lower concentrations? 

```{r, fig.height=6}
ufp0 %>%
  mutate(Study = paste0(Location, "; ", 
                        Study)
         ) %>%
   ggplot(aes(y= Study, #Location, 
             xmin = instrument_low_range_nm, xmax=instrument_high_range_nm,
             col = Mean,
             )) + 
  geom_linerange(size=1.5
    ) +
  scale_color_gradient(low = "yellow", high = "red") +
  
  scale_x_log10(breaks = c(4, 10, 20, 300, 1000, 3000)) +
  
  facet_grid(rows = vars(Sampling_type), 
             scales="free_y", #space="free_y"
             ) +
  
  labs(x = "Instrument Particle Size Range (nm)",
       col = "Avg PNC\n(pt/cm3)",
       title = "Particle size ranges collected in various PNC studies"
       )
  

```


 